### 2023-11-30 Data/DataPackageProtocol.Rmd
changed project structure based on personal template change.

# Provide here a full description of each dataset in the project not captured by
#  the automated dataset-specific codebooks.

# InputData --------------------------------------------------------------------

# IntermediateData -------------------------------------------------------------

# OutputData -------------------------------------------------------------------