### 2024-07-01 Scripts/Manuscript/AppendixScripts/Appendix_pre-reg_analyses.Rmd

<!-- Appendix A -->

# Pre-registration Analyses

This appendix contains the pre-registered analyses that were either changed or omitted from the article.

## Intraclasscorelation assessment for within measure variation in reliability

### Pre-registration Analysis Justification & Reason for Omission from Article
Reliabilities could be calculated from the replication datasets. On top of that, reliabilities could be calculated for each individual lab per replication set so that the variance of them could be assessed. To separate between and within variance in reliability, a multilevel random intercept model was specified with replication set as the grouping variable and the Cronbach's alpha or omega value of a replication as dependent using the lmer function from the lme4 package [@R-lme4] t.

```{r Hypothesis2Model, include = FALSE}
# intercept only model to check the within and between group variance.
H2_test_result_alpha <- lmer(formula = alpha ~ 1 + (1|g), data = data_h2)

```

The purpose of this analysis was to investigate variability in calculated reliability coefficients within and between replication sets. However, at the time of pre-registering we did not know of any formal test for the significance of within group variance after controlling for between group variance. Instead, the intraclass correlation (ICC) coefficient was calculated an indicator of within vs. between variability.

```{r ICCCalculations, include = FALSE}
# Variances
within_group_variance_alpha <- attr(VarCorr(H2_test_result_alpha), "sc")^2
between_group_variance_alpha <- attr(VarCorr(H2_test_result_alpha)$g, "stddev")^2

# ICC
ICC_alpha <- between_group_variance_alpha / (within_group_variance_alpha + between_group_variance_alpha)

```

This analysis was omitted from the article because it did not link well enough to hypothesis 2. The test results also do not give use a very informative representations of variance in reliability in the data for us to draw further inferences from. For instance, it assumes that we should observe large differences in reliabilities between measures for this to provide an accurate test of whether the variance within measures is large or not.

### Results
The between-group and within-group variance for Cronbach's alpha were approximately `r apa_num(between_group_variance_alpha, digits = 3)` and `r apa_num(within_group_variance_alpha, digits = 3)` respectively. The resulting ICC was approximately `r apa_num(ICC_alpha, digits = 3)`. This suggests that variance was larger between replication sets than within. These results are generally in line with hypothesis 2. However, this statistic does not necessarily give a test of hypothesis 2, as outlined above.

## Unidimensionality Exploratory Analyses

### Pre-registration Analysis Justification & Reason for Omission from Article

To present additional validity evidence alongside the analyses of Cronbach's alpha and omega coefficients, the unidimensionality of each measure was investigated. To test this, each multiple item scalar measure had their factor structure investigated on the entire measure data of the replication set.

This analysis was omitted from the main article since determining when a measure should be unidimensional would require an in-depth review and expert evaluation for the literature surrounding the measure, which is outside of the scope of this article and the authors' expertise. Additionally, this information may also simply not be available, further complicating the analyses.

For the sake of this analysis the simple heuristic was chosen that if the measure was used as a single composite variable in the analyses, then it should be uni-dimensional. This is somewhat similar to what was done by @shaw2020measurement, but for the sake of this article it was deemed to strong an assumption to make that this heuristic would always hold. For instance, IQ score is often used as a single indicator score, however the construct of IQ is not considered unidimensional, and is instead understood to have many sub-dimensions (e.g. verbal comprehension, perceptual reasoning, etc.). As a result, the use of a single score in the analyses cannot be taken as evidence that the construct is considered unidimensional by neither the authors, nor the broader literature on the topic.

Nevertheless, we present here the result of the analyses made under the assumption that the one score, one dimension heuristic holds, for the readers that are comfortable with this heuristic.

## Model

A single factor confirmatory factor analysis was implemented using the *fa* function from the *psych* package [@R-psych] in R. The maximum likelihood factoring method was used, with defaults being used for all other function arguments. From this analysis, the RMSEA of the one factor solution was extracted. Additionally, a parallel test was conducted using the *fa.parallel* function also from the psych package. If either the RMSEA score \< .08, or the parallel analysis returned a one factor solution, unidimensionality was coded as true, otherwise it was coded as false.

```{r UnidimensionalityExploratoryAnalysis, include = FALSE}
extracted_data_list <- list(data_1.10_clean, data_1.11_clean, data_1.12.3.1_clean, data_1.12.3.2_clean,data_2.12.1_clean, data_2.12.2_clean, data_2.12.3_clean, data_2.15_clean, data_2.20_clean, data_2.23_clean, data_3.2.1.1_clean, data_3.2.1.2_clean, data_3.7.1_clean, data_3.7.2_clean, data_3.8.2_clean, data_5.1.1_clean, data_5.1.2_clean, data_5.7_clean, data_5.9.1_clean)

# creating an empty dataframe to store the results in
cfa_results <- data.frame(RMSEA = 0,
                          N_factors = 0,
                          Single_factor = "")

# running the cfas for all of the extracted datasets
for (i in 1:length(extracted_data_list)){
  # obtaining single factor model cfa RMSEA value for evaluation
  RMSEA_value <- fa(extracted_data_list[[i]][-1])$RMSEA[1]

  if(is.null(RMSEA_value) != TRUE){
    RMSEA_value <- apa_num(RMSEA_value, digits = 3)
  }

  # conducting parallel test to check if one factor solution is best.
  parallel_N_factors <- fa.parallel(extracted_data_list[[i]][-1], 
                                    fa = "fa", plot = FALSE)$nfact

  # adding a single factor check indicator value
  single_factor <- ifelse(RMSEA_value < 0.08 | parallel_N_factors == 1, "Yes", 
                          "No")

  # adding the cfa results of one factor analysis to the results dataframe.
  cfa_results <- rbind(cfa_results, c(RMSEA_value, parallel_N_factors, 
                                      single_factor))
}

# removing the initial empty row
cfa_results <- cfa_results[-1,]

```

```{r cfaResultsStoring, include = FALSE, eval = FALSE}
# storing the cfa_results, so they can be referenced within the main manuscript.
open_data(data = cfa_results, filename = paste0(paste0(
  "Data/AnalysisData/", deparse(substitute(cfa_results))), ".csv"),
  codebook = paste0(paste0(
    "Data/AnalysisData/codebook_", 
    deparse(substitute(cfa_results))), ".Rmd"),
  value_labels = paste0(paste0(
    "Data/AnalysisData/value_labels_", 
    deparse(substitute(cfa_results))), ".yml")) 
```

## Results

Suitable data for factor analyses was available for `r apa_num(nrow(cfa_results), numerals = FALSE)` measures. These measures were the same as those shown in Figure \@ref(fig:Plot23AlphaCode). The results are shown in Table \@ref(tab:UnidimesnionalityTestResultsTable). A composite score based on the measurement responses was used for sixteen of the `r apa_num(nrow(cfa_results), numerals = FALSE)` measures to form a single variable in the analyses and were thus regarded as intended to be unidimensional. The number of dimensions for the remaining three were unclear. Sufficient fit of the unidimensional model was found for `r apa_num(sum(cfa_results$Single_factor == "Yes", na.rm = TRUE), numerals = FALSE)` out of the `r apa_num(nrow(cfa_results), numerals = FALSE)` evaluated measures. However, as shown in Table \@ref(tab:UnidimesnionalityTestResultsTable), based on the parallel analysis test alone the unidimensional model held for only `r apa_num(sum(cfa_results$N_factors == 1, na.rm = TRUE), numerals = FALSE)` of the `r apa_num(nrow(cfa_results), numerals = FALSE)` measures. For those `r apa_num(sum(cfa_results$N_factors == 1, na.rm = TRUE), numerals = FALSE)` measures and an additional `r apa_num(sum(cfa_results$RMSEA < .08, na.rm = TRUE) - sum(cfa_results$N_factors == 1, na.rm = TRUE), numerals = FALSE)` the RMSEA $< .08$ criteria suggested evidence for single factor structures.

```{r UnidimesnionalityTestResultsTable, warning = FALSE}
# add the names of the relevant papers and a shortcut for the measure used.
cfa_results_expanded <- cbind(c("Caruso et al. (2012)", "Husnu & Crisp (2010)", "Nosek et al. (2002), Art", "Nosek et al. (2002), Math", "Anderson et al. (2012), SWL", "Anderson et al. (2012), PA", "Anderson et al. (2012), NA", "Giessner & Schubert (2007)", "Norenzayan et al. (2002)", "Zhong & Liljenquist (2006)", "Monin & Miller (2001), most", "Monin & Miller (2001), some", "Cacioppo et al. (1983), elm", "Cacioppo et al. (1983), nfc", "De Fruyt et al. (2000), consc", "Albarracín et al. (2008), exp. 5, math", "Albarracín et al. (2008), exp. 5, verb",  "Shnabel & Nadler (2008)", "Vohs & Schooler (2008)"), cfa_results)

# adding sensible column names
colnames(cfa_results_expanded) <- c("Original Article", "RMSEA", "N Factors", "Unidimensional")

# Renaming the columns for the not converged cfa.
cfa_results_for_print <- cfa_results_expanded
cfa_results_for_print$RMSEA[15] <- "Did"
cfa_results_for_print$`N Factors`[15] <- "Not"
cfa_results_for_print$Unidimensional[15] <- "Converge"

# print the table in apa formatting
apa_table(
  cfa_results_for_print, align = c("l", "r", "r", "r")
  , caption = "RMSEA and Parallel Analysis Suggested Factors based on Factor Analysis
for Measure Data", note = "N Factors refer to the number of factors suggested based on parallel analysis of the measure data. The word or abbreviation after the author names and year indicates the specific measure within that original article that was analyed.", escape = FALSE, row.names = FALSE, placement= "htp")

```

These results are potentially quite worrying given that many studies made use of single item measures, which assume unidimensionality by default without being able to assess it. These results show that for the measures that could be tested for unidimensionality, and that were mostly used as single indicators in the analyses for the replication protocol, unidimensionality commonly did not hold. However, one could argue that measures being used as singular indicators does not necessarily mean these measures all assessed constructs with a unidimensional structure. For example, IQ is often used as a singular indicator, even though it consists of several factors [@weiss2013wisc]. However, it should be noted that at least with respect to Many Labs 2, all measures that were considered unidimensional for the analyses in this study were also considered unidimensional in the investigation by @shaw2020measurement who also observed a lack of unidimensionality evidence in their investigation of Many Labs 2. To conclude, there is an indication that validity evidence for measures used in replication research is lacking, and that this may have further implications beyond the measures for which it could be assessed here.

## H6 QMP-Category Specific Exploratory Analyses

### Pre-registration Analysis Justification & Reason for Omission from Article

Follow-up analyses were preregistered for hypothesis 6. Using a similar model to the main hypothesis test with all unique combination of the five different QMP types being tested across original and replication data. The goal of these analyses was to see if there are any type specific carry-over effects in QMPs between original and replication. Due to the large number of tests, the focus was exploratory using visualization rather than inference.

It is partly this large number of tests, as well as the resulting lack of clear interpretability in the results that caused this analysis to be omitted from the main article. However, for completeness and transparency it is reported here.

### Model

The model was the same as that used for the main test of hypothesis 6, only now QMP ratios for each of the five categories were used one at a time for each unique combination of original article and replication protocol QMP.

```{r Hypothesis6Associations, warning = FALSE}
## Associations between the QMP categories
# associations are generated using a helper function available in the source_ 
# script. These associations correspond to the coefficients between replication
# and original article QMPs of the various QMP types, as estimated through a 
# beta regression.
# Associations between the QMP categories (original QMP coding) MOVE THIS ANALYSIS TO INITIAL CODING APPENDIX
Results_H6_between <- Associations_H6(data_h6)

# keep in mind mod_ratio (not Rep) is based on a very small sample size, less
# of a concern for the revised.

# Associations between the QMP categories (revised QMP coding)
Results_H6_between_REV <- Associations_H6_REV(data_h6)

```

### Results

Planned exploratory analyses were preformed to expand on the test for hypothesis 6. The relationships between original article QMP ratios and replication protocol QMP ratios were investigated for each QMP type separately. Figure \@ref(fig:Plot46FullRevisedDataCode) visually illustrates the relations of the revised QMP ratios from different QMP types between replication protocols and original articles. As shown in the figure, no relation was strong nor highly stable, and few were consistent between initial and revised coding protocols.

```{r Plot46FullRevisedDataCode, error = FALSE, warning = FALSE, fig.cap = "Scatterplot of original and replication QMP ratios per type with linear regression line. The abbreviations on the axes relate to the QMP categories described in Table 1: Def. is definition, Op. is operationalization, Sel. is selection/creation, Quant. is quantification, and Mod. is modification. The X axis facets across the QMP ratios in the original articles, and the Y axis facets across the QMP ratios in the replication protocols. Each dot in the figure relates to describes the QMP ratio for the types in that graph across an original article and its replication protocol. Note: jitter was applied to these dots to show the number of observations at points where multiple dots were present"}
# Revised
ggplot(plot_46_data_rev, aes(QMP_ratio, QMP_Rep_ratio)) + 
  geom_point() +
  geom_jitter(width = 0.05, height = 0.05) +
  stat_smooth(method="lm") + 
  theme_bw() +
  theme(strip.text.y = element_text(size = 7), strip.background = element_rect(
    color = "black", fill = "white", size = 1, linetype = "solid"
  )) +
  xlab("QMP Ratio (Original)") + ylab("QMP Ratio (Replication)") +
  scale_x_continuous(labels = c("0", "0.25", "0.5", "0.75", "1")) + 
  facet_grid(vars(QMP_Rep_type), vars(QMP_type))

```

Table \@ref(tab:H6ExploratoryAssociationsRevisedTable) shows the beta-regression tests for each of the QMP category comparisons. The relations that are consistent are those between original article definition QMPs with replication selection and total QMPs, original article selection QMPs with replication total QMPs, and finally original article quantification QMPs with replication selection QMPs. All these relations were positive across both coding protocols (see Appendix D), illustrating that a greater QMP ratio of the specified type in an original article is associated with an increase in QMP ratio for the specific QMP type, or total QMP ratio in the replication protocol.

```{r H6ExploratoryAssociationsRevisedTable, warning = FALSE}
# creating the dataframe for the table
H6_table_dataframe_l_REV <- rbind(Results_H6_between_REV$def_def, Results_H6_between_REV$def_op,
  Results_H6_between_REV$def_sel, Results_H6_between_REV$def_quant, Results_H6_between_REV$def_mod, 
  Results_H6_between_REV$op_def, Results_H6_between_REV$op_op, Results_H6_between_REV$op_sel, 
  Results_H6_between_REV$op_quant, Results_H6_between_REV$op_mod, Results_H6_between_REV$sel_def, 
  Results_H6_between_REV$sel_op, Results_H6_between_REV$sel_sel, Results_H6_between_REV$sel_quant, 
  Results_H6_between_REV$sel_mod)

H6_table_dataframe_r_REV <- rbind(Results_H6_between_REV$quant_def, Results_H6_between_REV$quant_op,
  Results_H6_between_REV$quant_sel, Results_H6_between_REV$quant_quant, Results_H6_between_REV$quant_mod, 
  Results_H6_between_REV$mod_def, Results_H6_between_REV$mod_op, Results_H6_between_REV$mod_sel, 
  Results_H6_between_REV$mod_quant, Results_H6_between_REV$mod_mod, Results_H6_between_REV$tot_def, 
  Results_H6_between_REV$tot_op, Results_H6_between_REV$tot_sel, Results_H6_between_REV$tot_quant, 
  Results_H6_between_REV$tot_mod)

H6_table_dataframe_REV <- cbind(H6_table_dataframe_l_REV, H6_table_dataframe_r_REV)

# adding asterices for significance to the significant p in the dataframe
H6_table_dataframe_REV[11,3] <- paste0(H6_table_dataframe_REV[11,3], "$^*$")
H6_table_dataframe_REV[13,3] <- paste0(H6_table_dataframe_REV[13,3], "$^*$")
H6_table_dataframe_REV[14,3] <- paste0(H6_table_dataframe_REV[14,3], "$^*$")
H6_table_dataframe_REV[4,6] <- paste0(H6_table_dataframe_REV[4,6], "$^*$")
H6_table_dataframe_REV[11,6] <- paste0(H6_table_dataframe_REV[11,6], "$^*$")
H6_table_dataframe_REV[13,6] <- paste0(H6_table_dataframe_REV[13,6], "$^*$")
H6_table_dataframe_REV[14,6] <- paste0(H6_table_dataframe_REV[14,6], "$^*$")

# adding a column to help indicate the original QMP type
H6_table_dataframe_REV_expanded <- cbind(c("Definition", "Operationalisation", "Selection", "Quantification", "Modification", "Definition", "Operationalisation", "Selection", "Quantification", "Modification", "Definition", "Operationalisation", "Selection", "Quantification", "Modification"), H6_table_dataframe_REV)

# adding rows to give indication of specific sections (apa_table being awkward)
H6_table_REV_for_printing <- rbind(c("QMP Type", "", "Definition", "", "", "Quantification", ""), H6_table_dataframe_REV_expanded[1:5,], c("QMP Type", "", "Operationalisation", "", "", "Modification", ""),  H6_table_dataframe_REV_expanded[(6:10),], c("QMP Type", "", "Selection", "", "", "Total", ""), H6_table_dataframe_REV_expanded[(11:15),])

colnames(H6_table_REV_for_printing) <-  c("Original", "$b$", "SD", "$p$", "$b$", "SD", "$p$")

# print the table in apa formatting
apa_table(
  H6_table_REV_for_printing, align = c("l", "r", "c", "r", "r", "c", "r")
  , caption = "Beta Regression Test Results for the Association between Replication
QMP Ratio and Original QMP Ratio as obtained using the Revised Protocol"
  , note = "Rows indicate Original QMP type, and within table indicates Replication QMP type. Sample size for comparisons involving modification were lower since not all measures were modified. Therefore, the standard errors are larger. $^* p < .05$"
  , escape = FALSE, placement= "htp", midrules = c(1, 6, 7, 12, 13))

```

A possible explanation for these results could be that original articles have to provide a clear definition as well as details on the scale, in order for a replication to be able to justify selecting the measure. A replication may also be in general more dependent on the original article for information on the definition of the constructs and the choice of measurement, compared to a description of the operationalisation, quantification, and any modifications. These former are features of a measurement, which while preferably similar in a replication to the original, can and should be reported in the replication protocol even if the original article does not contain the details on these features itself. However, these are just speculations as the results showed no clear relations between different QMP types across original and replication (see Figure \@ref(fig:Plot46FullRevisedDataCode)), the number of tests were many while multiple testing was not corrected for, and finally testing of causal links was not possible with this data.