### 2024-07-01 SupplementaryMaterials/data_retrieval_information.Rmd

# Data retrieval information

This supplementary material file contains information on the search strategy that was used to retrieve the three main data sources used in this study: the replication protocols, replication datasets, and the original articles.

Furthermore, for the replications protocols and replication datasets, we specify here the exact OSF file paths and file names of the specific protocols and datasets that were retrieved.

## Replication Protocols

### Search

The OSF page for each Many Labs project was combed through, in order to identify the documents that contained information on the measurement practice and reported reliability. For Many Labs 1 2, & 3 a single replication protocol for all studies could be identified, while for Many Labs 5 protocols were available for each set of replications separately. For Many Labs 5 for each replication set the relevant OSF page was searched through for a protocol file.

If available, a version of the protocol labelled as revised, post-review, peer-review, or endorsed was selected. Otherwise, the latest uploaded relevant protocol was selected. Protocols labelled as data collection, analysis protocol or anything similar were excluded.

The steps taken to access the replication protocol data from Many Labs 5 were not preregistered, and only became clear upon further investigation of the accessible information.

### OSF pages

Below are the relevant OSF pages, and exact path and file names of the relevant pre-registered replication protocols that were extracted, to be then coded on measurement reporting information in the Many Labs replications projects.

Many Labs 1 (<https://osf.io/wx7ck/>):
- Path: OSF Storage (United States)/
- Protocol File: ManyLabsProposal_V1.1.docx

Many Labs 2 (<https://osf.io/8cd4r/>):
- Path: OSF Storage (United States)/
- Protocol File: ML2_-_Protocol.pdf

Many Labs 3 (<https://osf.io/ct89g/>):
- Path: OSF Storage (United States)/
- Protocol File: ML3_Protocol_9152014.pdf

Many Labs 5 Study 1, Albarracín et al. (2008), Experiment 5 (<https://osf.io/a9hk2/>):
-	Path: ML5 Site Replications/Protocol and Analysis Plan/OSF Storage (United States)/
-	Protocol File: ML5Alb5Pre-RegProtocolAndAnalysisPlan.docx

Many Labs 5 Study 2, Albarracín et al. (2008), Experiment 7 (<https://osf.io/j5s8c/>):
-	Path: OSF Storage (United States)/Proposal and Manuscript/
-	Protocol File: Albarracinetal.Study7RevisedProtocol.docx

Many Labs 5 Study 3, Förster, Liberman, & Kuschel (2008) (<https://osf.io/rje4t/>):
-	Path: Materials/Revised Protocol/
-	Protocol File: Förster et al. Replication Endorsed Protocol Final.docx

Many Labs 5 Study 4, LoBue & DeLoache (2008) (<https://osf.io/9g7pw/>):
-	Path: OSF Storage (United States)/ 0. Protocols/
-	Protocol File: LoBueDeLoache2008_ML5RevisedProtocol.pdf

Many Labs 5 Study 5, Payne, Burkley, & Stokes (2008) (<https://osf.io/wxd4g/>):
-	Path: OSF Storage (United States)/
-	Protocol File: PayneBurkleyStokes2008RevisedProtocol.docx

Many Labs 5 Study 6, Risen & Gilovich (2008) (<https://osf.io/h5a9y/>):
-	Path: Dropbox: ML5_risen_Gilovich_git/protocols/Post-review protocol (for preregistration)/
-	Protocol File: Risen & Gilovich Revised Protocol.docx

Many Labs 5 Study 7, Shnabel & Nadler (2008) (<https://osf.io/vz5ru/>):
-	Path: OSF Storage (United States)/
-	Protocol File: ML5 ShnabelNadler Revised Protocol.docx

Many Labs 5 Study 8, van Dijk et al. (2008)) (<https://osf.io/f6n27/>):
-	Path: Replications/OSF Storage (United States)/Protocols/
-	Protocol File: vanDijketal.2008RevisedProtocol.pdf

Many Labs 5 Study 9, Vohs & Schooler (2008) (<https://osf.io/8rcbk/>):
-	Path: Research Protocols (and other things you might want to read)/OSF Storage (United States)/
-	Protocol File: ML5 Protocol.docx

## Registered Replication Datasets
Below are the relevant OSF pages, and exact path and file names of the relevant replication datasets that were extracted, to be then used for calculating the reliability and testing the unidimensionality of the measures used in the Many Labs replications projects.

Many Labs 1 (<https://osf.io/wx7ck/>):
-	Path: Data/Datasets.zip
-	Data File: Full_Dataset_De-Identified.sav

Many Labs 2 Study 2, Kay et al. (2014) (<https://osf.io/8cd4r/>):
-	Path: Data and Results (per site, etc.)/OSF Storage (United States)/Structure & Goal Pursuit (Kay et al., 2014)/Kay.1/Global/Data/
-	Data File: Kay_1_study_global_include_all_CLEAN_CASE.csv

Many Labs 2 Study 3, Alter et al. (2007) (<https://osf.io/8cd4r/>):
-	Path: Data and Results (per site, etc.)/OSF Storage (United States)/Incidental Disfluency (Alter et al., 2007)/Alter.1/Global/Data/
-	Data File: Alter_1_study_global_include_all_CLEAN_CASE.csv

Many Labs 2 Study 4 measure 1, Graham et al. (2009) (<https://osf.io/8cd4r/>):
-	Path: Data and Results (per site, etc.)/OSF Storage (United States)/Moral Foundations (Graham et al., 2009)/Graham.1/Global/Data/
-	Data File: Graham_1_study_global_include_all_CLEAN_CASE.csv

Many Labs 2 Study 4 measure 2, Graham et al. (2009) (<https://osf.io/8cd4r/>):
-	Path: Data and Results (per site, etc.)/OSF Storage (United States)/Moral Foundations (Graham et al., 2009)/Graham.2/Global/Data/
-	Data File: Graham_2_study_global_include_all_CLEAN_CASE.csv

Many Labs 2 Study 8 measure 2, Inbar et al. (2009) (<https://osf.io/8cd4r/>):
-	Path: Data and Results (per site, etc.)/OSF Storage (United States)/Disgust & Homophobia (Inbar et al., 2009)/Graham.1b/Global/Data/
-	Data File: Inbar_1b_study_global_include_all_CLEAN_CASE.csv

Many Labs 2 Study 10 measure 1, vanLange et al. (1997) (<https://osf.io/8cd4r/>):
-	Path: Data and Results (per site, etc.)/OSF Storage (United States)/Social Value Orientation (Van Lange et al., 1997)/vanLange.1/Global/Data/
-	Data File: vanLange_1_study_global_include_all_CLEAN_CASE.csv

Many Labs 2 Study 12, Anderson et al. (2012) (<https://osf.io/8cd4r/>):
-	Path: Data and Results (per site, etc.)/OSF Storage (United States)/SMS & Well-Being (Anderson et al., 2012)/Anderson.1/Global/Data/
-	Data File: Anderson_1_study_global_include_all_CLEAN_CASE.csv

Many Labs 2 Study 15, Giessner & Schubert (2007) (<https://osf.io/8cd4r/>):
-	Path: Data and Results (per site, etc.)/OSF Storage (United States)/Position & Power (Giessner & Schubert, 2007)/Giessner.1/Global/Data/
-	Data File: Giessner_1_study_global_include_all_CLEAN_CASE.csv

Many Labs 2 Study 19 measure 1, Savani et al. (2010) (<https://osf.io/8cd4r/>):
-	Path: Data and Results (per site, etc.)/OSF Storage (United States)/Actions are Choices (Savani et al., 2010)/Savani.3a/Global/Data/
-	Data File: Savani_3a_study_global_include_all_CLEAN_CASE.csv

Many Labs 2 Study 20, Norenzayan (2002) (<https://osf.io/8cd4r/>):
-	Path: Data and Results (per site, etc.)/OSF Storage (United States)/Intuitive Reasoning (Norenzayan et al., 2002)/Savani.1/Global/Data/
-	Data File: Norenzayan_1_study_global_include_all_CLEAN_CASE.csv

Many Labs 2 Study 23, Zhong & Liljenquist (2006) (<https://osf.io/8cd4r/>):
-	Path: Data and Results (per site, etc.)/OSF Storage (United States)/Moral Cleansing (Zhong & Liljenquist, 2006)/Zhong.1/Global/Data/
-	Data File: Zhong_1_study_global_include_all_CLEAN_CASE.csv

Many Labs 3 (<https://osf.io/ct89g/>):
-	Path: OSF Storage (United States)/ML3 Final Data.zip
-	Data File: ML3AllSitesandmTurk.csv

Many Labs 5 Study 1, Albarracín et al. (2008), Experiment 5 (<https://osf.io/a9hk2/>):
-	Path: Data/OSF Storage (United States)/
-	Data File: ML5 Alb 5 Revised Protocol In Lab.xlsx

Many Labs 5 Study 4, LoBue & DeLoache (2008) (<https://osf.io/9g7pw/>):
-	Path: ML5 Dataset & Analysis LoBue & DeLoache/OSF Storage (United States)/
-	Data File: dataset_LoBue_raw.sav

Many Labs 5 Study 5, Payne, Burkley, & Stokes (2008) (<https://osf.io/wxd4g/>):
-	Path: Data/OSF Storage (United States)/
-	Data File: Genova_ML5.csv

Many Labs 5 Study 7, Shnabel & Nadler (2008) (<https://osf.io/vz5ru/>):
-	Path: Data, Materials, Reports: ML5 Replication of Shnabel & Nadler (2008)/Data, Codebook, and syntax/OSF Storage (United States)/
-	Data File: ml5_computed_dataset.csv

Many Labs 5 Study 9 measure 1, Vohs & Schooler (2008) (<https://osf.io/8rcbk/>):
-	Path: Data/OSF Storage (United States)/FinalData/
-	Data File: ML5_fulldataset_with_suspicion.csv

## Original Articles
The original articles were identified using the citations for these articles in each replication protocol. They were then retrieved and extracted through searches on Web of Science, Google Scholar, and PsycINFO (in that order). All articles except Hyman and Sheatsley (1950) could be retrieved through this method. Howeever, after contacting the first author of Klein et al. (2014) a suitable version of the article was obtained. Thus in the end all original articles could be retrieved.

# Plot of data information

This can only be run after data is loaded with the main manuscript file.

```{r DataStructurePlotCode, warning = FALSE, fig.cap = "This figure displays the sample size for the included original articles and total sample size for multiple replications. The color of each point indicates whether it represents an original article or a replication. Each point's shape relates to whether the hypothesis was supported or not, while the size indicates the number of measures used within that article or replication. The figure is log10-scaled, each light grey vertical line represents a step of 500 in sample size. Two original articles are not represented in this figure due to them containing no reported sample size.", fig.height = 4, fig.width = 12}

# plot of the multilevel data structure
ggplot(plot_data_structure_data, aes(x = N, y = factor(many_labs_version, levels = c("5", "3", "2", "1")), 
                                colour = rep_org, shape = hypothesis_support, size = measure_count)) +
  geom_point(alpha = .33) +

  scale_x_continuous(trans = "log10", breaks = c(10, 100, 1000, 10000), 
                     labels = c(10, 100, 1000, 10000), limits = c(10, 10000),  
                     minor_breaks = seq(500, 10000, 500)) +

  theme_minimal() +
  theme(plot.margin = unit(c(1, 6.5, 1, 1), "lines"), 
        panel.grid.major.x = element_line(color = "grey",
                                          size = 0.5,
                                          linetype = 1),
        text = element_text(size = 15)) +

  guides(color = guide_legend(nrow = 1, byrow = TRUE)) +
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  guides(size = guide_legend(nrow = 2, byrow = TRUE)) +

  labs(x = "N", y = "Many Labs Version", color = "Obtained from:", 
       size = "Nr. measures", shape = "Replicated")

```