### 2024-08-29 Data/Data_package_description.Rmd

<!-- Provide here a full description of each dataset in the project not captured by the automated dataset-specific codebooks. -->

# InputData

### Coded_Data_Initial_raw.csv

*Origin*: This dataset was created originally for this research project. It was obtained using the Measurement_Error_Reporting_Coding_Protocol.pdf that can be found in the SupplementaryMaterials/CodingProtocols folder of this project on the Many Labs protocols mentioned in the texts and the related original articles for each of the many labs replications.

*Explanation*: The information in this dataset was used for analyses relating to QMP ratios among original articles and replication studies. It additionally contains peripheral information about the articles and study in it that were used throughout the analyses and for descriptives. Questions related to modification had 'no modification' as an additional answer option, in case that specific aspect of the measure was explicitly not modified, but might possibly have been applicable, because it was a relevant aspect albeit unmodified. But these operated equivalently to the Not applicable response with respect to the calculation of the QMP ratio. They are included here for future reference.

*Variables*:

-   **StartDate, EndData, Status, IPAdress, Progress, Duration..in.seconds., Finished, RecordedDate, Responseid, RecipientLastName, RecipientFirstName, RecipientEmail, ExternalReference, LocationLatitude, LocationLongitude, DistributionChannel,** & **UserLanguage:** automated Qualtrics metadata, not used in study.

-   **Many_labs_version:** Indicates using a number the Many Labs version the study/replication under investigation belongs to.

-   **research_type:** Indicates using text whether a coded measurement comes from a Replication protocol, or an Original article.

-   **research_title:** the title of the replicated study as it appears in the associated Many Labs protocol/original article.

-   **measure_name:** the measure's name as stated within the Many Labs protocol/original article, or a short description based on what was said in the many labs protocol/original article.

-   **variable_name:** the name of the variable the measure was intended to measure based on the description in the many labs protocol/original article.

-   **Q58:** Boolean variable indicating whether or not the variable assessed by the measure was assessed as [part of a larger measure intending to measure multiple variables/dimensions of a construct.

-   **variable_order:** a number indicating the order of appearance for the measure within the total number of measures for that study/article.

-   **N:** the number of people that took the measure as indicated in the many labs report/original article.

-   **N_items:** a three choice option variable indicating whether a measure was single item, multiple item, or not a scale expected to have regular psychometric properties.

-   **hypothesis_support:** question indicating whether or not the Many Labs replications found support for the original hypothesis at a meta-analytic alpha level of .05 in the same direction as the original effect.

-   Reliability Coefficient Variables

    -   **reliability_type:** for multiple item measures only, a question asking if a reliability coefficient was reported, and if so which one with a couple of options and the option to select Other.

    -   **reliability_type_5_TEXT:** for questions that answered Oher on the reliability_type question, a text description of the specific reliability index.

    -   **reliability_coeff:** for all measures with a reliability coefficient, the value of that coefficient as reported in the Many Labs protocol or original article.

-   QMP Variables

    -   **Def1:** A boolean variable with an additional Not Applicable option, indicating whether or not a psychological/sociological definition was given within the protocol study description or original article to the measured variable from the measure.

    -   **Op_version:** if described within the protocol or article, the name and citation for the specific version of the measure that was used.

    -   **Op1:** A boolean variable with an additional Not Applicable option, indicating if the specific version of the measure is described.

    -   **Op2:** A boolean variable with an additional Not Applicable option, indicating whether or not the administration format of the measure and the environment the measure was taken in was described.

    -   **Op3:** A boolean variable with an additional Not Applicable option, indicating whether or not the steps taken in administering the measure were described or not.

    -   **Op4:** A boolean variable with an additional Not Applicable option, indicating whether or not the authors justified the way they conducted the measurement and why the specific choice was necessary to properly measure what they were investigating.

    -   **Op5:** A boolean variable with an additional Not Applicable option, indicating whether or not example items are presented in the text, or in an appendix.

    -   **Sel_existing:** Variable indicating whether existing articles were used, the measure was newly created for the study, or if it was not clearly stated.

    -   **Sel_existing_2_TEXT:** A textual description giving the (general) name or explanation of which previous measure was used, in case an already existing measure was used.

    -   **Sel1:** A boolean variable with an additional Not Applicable option, indicating whether or not the source of the scale is provided, including if it is new that this is clearly stated.

    -   **Sel2:** A boolean variable with an additional Not Applicable option, indicating whether or not reasoning is given for choosing a particular measure either existing or creating a new one.

    -   **Sel3:** A boolean variable with an additional Not Applicable option, indicating whether or not psychometric evidence for the validity of a scale was presented.

    -   **Sel4:** A boolean variable with an additional Not Applicable option, indicating whether or not there is psychometric evidence for the measure outside of the main analyses of the protocol or study, this can be either from an earlier study, or from a pilot study testing the measure.

    -   **Sel_psych_evidence:** A variable meant to indicate whether an analysis of the measure's validity was applicable, and if so whether or not it was there, and if so too, what type of psychometric evidence was reported.

    -   **Sel_psych_evidence_4_TEXT:** A textual description of the specific analyses of validity that was used in case the Other, namely: option was picked in the last question.

    -   **Quant1:** A boolean variable with an additional Not Applicable option, indicating whether or not the number of items in the measure are described.

    -   **Quant2:** A boolean variable with an additional Not Applicable option, indicating whether or not the number of response options for the items in the measure are described.

    -   **Quant3:** A boolean variable with an additional Not Applicable option, indicating whether or not recoding of responses is indicated for multiple item measures.

    -   **Quant4:** A boolean variable with an additional Not Applicable option, indicating whether or not the aggregation of the scores on the items on a multiple item measure into a single index was described.

    -   **Mod_check:** A variable indicating whether or not measurement modifications were mentioned at all, and if so whether the measurement was modified or not.

    -   **Mod1:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect) and Not Applicable option, indicating whether or not any modification to the administration format was mentioned or not.

    -   **Mod2:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect) and Not Applicable option, indicating whether or not there is validity evidence to support the change in the administration format.

    -   **Mod3:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect) and Not Applicable option, indicating whether or not any modification to the language of the questionnaire was mentioned or not.

    -   **Mod4:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect) and Not Applicable option, indicating whether or not there is validity evidence for the translated measure.

    -   **Mod5:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect) and Not Applicable option, indicating whether or not any a change in the number of items or response options deviated from the orginal source.

    -   **Mod6:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect) and Not Applicable option, indicating whether or not there is validity evidence for the measure with a change in the number of items or response options.

    -   **Mod_time:** A variable indicating whether or not the moment the modifications was reported, and if so whether this occurred before or after the data collection for the study.

### Coded_Data_Revised_raw.csv

*Origin*: This dataset was created originally for this research project. It was obtained using the Measurement_Error_Reporting_Revised_Coding_Protocol.pdf that can be found in the SupplementaryMaterials/CodingProtocols folder of this project on the Many Labs protocols mentioned in the texts and the related original articles for each of the many labs replications (the same sample as was used to obtain the Coded_Data_Initial_raw.csv).

*Explanation*: After discussion with the supervisor based on the experiences of the initial coding round it was determined that several of the demarcation for meeting several QMP criteria were likely too strict, and thus several of the more strictly coded questions were revised and coded again using a new coding protocol. The goal being to offer an as lenient as possible overview of the errors in reporting in the investigated literature.

*Variables*:

-   **StartDate, EndData, Status, IPAdress, Progress, Duration..in.seconds., Finished, RecordedDate, Responseid, RecipientLastName, RecipientFirstName, RecipientEmail, ExternalReference, LocationLatitude, LocationLongitude, DistributionChannel,** & **UserLanguage:** automated Qualtrics metadata, not used in study.

-   **Many_labs_version:** Indicates using a number the Many Labs version the study/replication under investigation belongs to.

-   **research_type:** Indicates using text whether a coded measurement comes from a Replication protocol, or an Original article.

-   **research_title:** the title of the replicated study as it appears in the associated Many Labs protocol/original article.

-   **variable_order:** a number indicating the order of appearance for the measure within the total number of measures for that study/article.

-   QMP Variables

    -   **Op1:** A boolean variable with an additional Not Applicable option, indicating if the specific version of the measure is described, or the item in a single item measure is itself described in sufficient detail.

    -   **Op2:** A boolean variable with an additional Not Applicable option, indicating whether or not the administration format of the measure and the environment the measure was taken in was described generally in as far as it is relevant to the construct.

    -   **Op5:** A boolean variable with an additional Not Applicable option, indicating whether or not example items are presented in the text, or in an (online) appendix.

    -   **Sel1:** A boolean variable with an additional Not Applicable option, indicating whether or not the source of the scale is provided. In case the scale is new, the context should make this apparent such that no source needs to be provided for this scale.

    -   **Sel3:** A boolean variable with an additional Not Applicable option, indicating whether or not psychometric evidence for the validity of a scale was presented. Or in case such evidence is not relevant for that particular measure, a statement that clarifies this.

    -   **Sel_psych_evidence:** A variable meant to indicate whether an analysis of the measure's validity was applicable, and if so whether or not it was there, and if so too, what type of psychometric evidence was reported.

    -   **Sel_psych_evidence_4_TEXT:** A textual description of the specific analyses of validity that was used in case the Other, namely: option was picked in the last question.

    -   **Quant1:** A boolean variable with an additional Not Applicable option, indicating whether or not the number of items in the measure are described.

    -   **Quant2:** A boolean variable with an additional Not Applicable option, indicating whether or not the number of response options for the items in the measure are described.

    -   **Quant3:** A boolean variable with an additional Not Applicable option, indicating whether or not recoding of responses is indicated for multiple item measures.

    -   **Mod_check:** A variable indicating whether or not there were known modifications to the measurement before use in the research, and if so whether the measurement was modified or not.

    -   **Mod1:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect) and Not Applicable option, indicating whether or not any modification to the administration format of the overarching study protocol was mentioned or not.

    -   **Mod2:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect, or the modification was of little impact) and Not Applicable option, indicating whether or not there is validity evidence to support the change in the administration format.

    -   **Mod3:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect) and Not Applicable option, indicating whether or not any modification to the language of the overarching study protocol was mentioned or not.

    -   **Mod4:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect, or the modification was of little impact) and Not Applicable option, indicating whether or not there is validity evidence for the translated measure.

    -   **Mod5:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect) and Not Applicable option, indicating whether or not any a change in the number of items or response options deviated from the original source, for those items relevant to the original source.

    -   **Mod6:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect, or the modification was of little impact) and Not Applicable option, indicating whether or not there is validity evidence for the measure with a change in the number of items or response options.

### Coded_Data_Vignette_raw.csv

*Origin*: This dataset was created originally for this research project. It was obtained using the Vignette_Item_Coding_Protocol.pdf that can be found in the SupplementaryMaterials/CodingProtocols folder of this project on the Many Labs protocols mentioned in the texts and the related original articles for each of the many labs replications (the same sample as was used to obtain the Coded_Data_Initial_raw.csv).

*Explanation*: This dataset was collected after the initial round of coding with the intention to control on certain analyses whether items described a vignette and not an ordinary questionnaire, which might have different standards with regards to measurement reporting and reliability. As of yet, it has remained unused in any analyses

*Variables*:

-   **RecordedDate**: Records the date the variable was coded, was not used further.

-   **Vignette**: Boolean variable indicating whether or not a particular measure was inseparably tied to a vignette describing some form of quoted text, visual, imagined scenario, cognitive task, or game.

### Data_ML1.csv

*Origin*: This dataset was obtained from the Datasets.zip archive as "Full_Dataset_De-Identified.sav", from the following OSF page: <https://osf.io/pqf9r/>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Ratliff, K., Vianello, M., Adams Jr, R. B., Bahník, S., Bernstein, M. J., . . . Nosek, B. A. (2014). Investigating variation in replicability: A "many labs" replication project. Social Psychology, 45 (3), 142--152. <https://doi.org/10.1027/1864-9335/a000178>, the dataset itself has the following doi: 10.17605/OSF.IO/PQF9R.

*Alterations*: the only alterations made to the dataset at this stage were: the name of the dataset was changed, and the format was changed from .sav to .csv.

### Data_2.2.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "Structure & Goal Pursuit (Kay et al., 2014)/Kay.1/Global/Data/" as "Kay_1_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.3.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "Incidental Disfluency (Alter et al., 2007)/Alter.1/Global/Data/" as "Alter_1_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.4.1.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "Moral Foundations (Graham et al., 2009)/Graham.1/Global/Data/" as "Graham_1_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.4.2.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "Moral Foundations (Graham et al., 2009)/Graham.2/Global/Data/" as "Graham_2_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.8.2.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "Disgust & Homophobia (Inbar et al., 2009)/Inbar.1b/Global/Data/" as "Inbar_1b_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.10.1.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "Social Value Orientation (Van Lange et al., 1997)/vanLange.1/Global/Data/" as "vanLange_1_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.12.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "SMS & Well-Being (Anderson et al., 2012)/Anderson.1/Global/Data/" as "Anderson_1_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.15.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "Position & Power (Giessner & Schubert, 2007)/Giessner.1/Global/Data/" as "Giessner_1_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.19.1.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "Actions are Choices (Savani et al., 2010)/Savani.3a/Global/Data/" as "Savani_3a_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.20.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "Intuitive Reasoning (Norenzayan et al., 2002)/Norenzayan.1/Global/Data" as "Norenzayan_1_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.23.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "/Moral Cleansing (Zhong & Liljenquist, 2006)/Zhong.1/Global/Data/" as "Zhong_1_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_ML3.csv

*Origin*: This dataset was obtained from the ML3 Final Data.zip archive in the base folder as "ML3AllSitesandmTurk.csv", from the following OSF page: <https://osf.io/bxw8j>.

*Attribution*: The datatset is related to the following article: Ebersole, C. R., Atherton, O. E., Belanger, A. L., Skulborstad, H. M., Allen, J. M., Banks, J. B., . . . Nosek, B. A. (2016). Many labs 3: Evaluating participant pool quality across the academic semester via replication. Journal of Experimental Social Psychology, 67, 68--82. <https://doi.org/10.1016/j.jesp.2015.10.012>.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_3.5.csv

*Origin*: This dataset was obtained from the ML3 Final Data.zip archive in the base folder as "ML3StroopData.csv", from the following OSF page: <https://osf.io/bxw8j>.

*Attribution*: The datatset is related to the following article: Ebersole, C. R., Atherton, O. E., Belanger, A. L., Skulborstad, H. M., Allen, J. M., Banks, J. B., . . . Nosek, B. A. (2016). Many labs 3: Evaluating participant pool quality across the academic semester via replication. Journal of Experimental Social Psychology, 67, 68--82. <https://doi.org/10.1016/j.jesp.2015.10.012>.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_5.1.csv

*Origin*: This dataset was obtained as "ML5 Alb 5 Revised Protocol In Lab.xlsx", from the following OSF page: <https://osf.io/64njs>.

*Attribution*: The datatset is related to the following article: Ebersole, C. R., Mathur, M. B., Baranski, E., Bart-Plange, D.-J., Buttrick, N. R., Chartier, C. R., . . . Nosek, B. A. (2020). Many labs 5: Testing pre-data-collection peer review as an intervention to increase replicability. Advances in Methods and Practices in Psychological Science, 3 (3), 309--331. <https://doi.org/10.1177/2515245920958687>.

*Alterations*: the only alterations made to the dataset at this stage were: the name of the dataset was changed, and the format was changed from .xlsx to .csv.

### Data_5.4.csv

*Origin*: This dataset was obtained as "dataset_LoBue_raw.sav", from the following OSF page: <https://osf.io/w7ft9>.

*Attribution*: The datatset is related to the following article: Ebersole, C. R., Mathur, M. B., Baranski, E., Bart-Plange, D.-J., Buttrick, N. R., Chartier, C. R., . . . Nosek, B. A. (2020). Many labs 5: Testing pre-data-collection peer review as an intervention to increase replicability. Advances in Methods and Practices in Psychological Science, 3 (3), 309--331. <https://doi.org/10.1177/2515245920958687>.

*Alterations*: the only alterations made to the dataset at this stage were: the name of the dataset was changed, and the format was changed from .sav to .csv.

### Data_5.5.csv

*Origin*: This dataset was obtained as "Genova_ML5.csv", from the following OSF page: <https://osf.io/prj7y>.

*Attribution*: The datatset is related to the following article: Ebersole, C. R., Mathur, M. B., Baranski, E., Bart-Plange, D.-J., Buttrick, N. R., Chartier, C. R., . . . Nosek, B. A. (2020). Many labs 5: Testing pre-data-collection peer review as an intervention to increase replicability. Advances in Methods and Practices in Psychological Science, 3 (3), 309--331. <https://doi.org/10.1177/2515245920958687>.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_5.7.csv

*Origin*: This dataset was obtained as "ml5_computed_dataset.csv", from the following OSF page: <https://osf.io/da8bh>.

*Attribution*: The datatset is related to the following article: Ebersole, C. R., Mathur, M. B., Baranski, E., Bart-Plange, D.-J., Buttrick, N. R., Chartier, C. R., . . . Nosek, B. A. (2020). Many labs 5: Testing pre-data-collection peer review as an intervention to increase replicability. Advances in Methods and Practices in Psychological Science, 3 (3), 309--331. <https://doi.org/10.1177/2515245920958687>.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_5.9.csv

*Origin*: This dataset was obtained as "ML5_fulldataset_with_suspicion.csv", from the following OSF page: <https://osf.io/5w6e7>.

*Attribution*: The datatset is related to the following article: Ebersole, C. R., Mathur, M. B., Baranski, E., Bart-Plange, D.-J., Buttrick, N. R., Chartier, C. R., . . . Nosek, B. A. (2020). Many labs 5: Testing pre-data-collection peer review as an intervention to increase replicability. Advances in Methods and Practices in Psychological Science, 3 (3), 309--331. <https://doi.org/10.1177/2515245920958687>.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

<!-- Should look into the datasets that were considered not usable and see what I might find for them now. Particularly Data_2.4.1, Data_2.4.2, Data_5.5, and Data_2.19.2 -->

# IntermediateData

### coded_data_original.csv

*Explanation*: contains all coded data on the original articles with renamed columns, fixed bad data, restructured to match the ordering of the protocols more eligibly, recoded to be close to suitable for analysis, and with several composite variables added that were calculated from the raw data.

*Newly Coded Variables*:

-   **\_REV variables:** these relate to the coding obtained for that particular variable using the revised coding protocol instead of the initial coding protocol.

-   **Inseparable material:** the Vignette variable from the Coded_Data_Vignette_raw.csv dataset.

-   **\_count variables**: a sum of the total number of TRUE coded variables within the named QMP category or the total across all categories.

-   **\_total variables**: the sum of all the variables within the named QMP category or the total across all categories that was not coded as Not Applicable or No Modification.

-   **\_ratio variables**: the ratio (rounded to four decimals) between the count of TRUE scores and the total applicable variables within the named QMP category or the total across all categories.

### coded_data_original_shortened.csv

*Explanation*: similar to the coded_data_original.csv only with the coded data on 3 measures that did not appear in the replication protocols being removed to allow for easier comparisons between replications and original studies.

### coded_data_replications.csv

*Explanation*: contains all coded data on the replications as described in the many labs protocols with renamed columns, fixed bad data, restructured to match the ordering of the protocols more ineligibly, recoded to be close to suitable for analysis, and with several composite variables added that were calculated from the raw data.

*Newly Coded Variables*:

-   **\_REV variables:** these relate to the coding obtained for that particular variable using the revised coding protocol instead of the initial coding protocol.

-   **Inseparable material:** the Vignette variable from the Coded_Data_Vignette_raw.csv dataset.

-   **\_count variables**: a sum of the total number of TRUE coded variables within the named QMP category or the total across all categories.

-   **\_total variables**: the sum of all the variables within the named QMP category or the total across all categories that was not coded as Not Applicable or No Modification.

-   **\_ratio variables**: the ratio (rounded to four decimals) between the count of TRUE scores and the total applicable variables within the named QMP category or the total across all categories.

### Data_1.3_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to anchoring used in study 3 (as stated in the Many Labs 1 Proposal_V1.1) from Many Labs 1 that was checked for reliability and unidimensionality.

### Data_1.10_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to system justification used in study 10 (as stated in the Many Labs 1 Proposal_V1.1) from Many Labs 1 that was checked for reliability and unidimensionality.

### Data_1.11_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to intentions to engage used in study 11 (as stated in the Many Labs 1 Proposal_V1.1) from Many Labs 1 that was checked for reliability and unidimensionality.

### Data_1.12.3.1_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to explicit attitudes towards art used in study 12 (as stated in the Many Labs 1 Proposal_V1.1) from Many Labs 1 that was checked for reliability and unidimensionality.

### Data_1.12.3.2_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to explicit attitudes towards math used in study 12 (as stated in the Many Labs 1 Proposal_V1.1) from Many Labs 1 that was checked for reliability and unidimensionality.

### Data_2.10.1_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to social value orientation used in study 10 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

### Data_2.12.1_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to satisfaction with life used in study 12 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

### Data_2.12.2_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to positive affect used in study 12 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

### Data_2.12.3_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to negative affect used in study 12 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

### Data_2.15_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to rated power used in study 15 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

### Data_2.20_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to rule or similarity based judgement used in study 20 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

### Data_2.23_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to cleaning product ratings used in study 23 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

### Data_3.2.1.1_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to list of most expressions used in study 2 (as stated in the Many Labs 3 protocol) from Many Labs 3 that was checked for reliability and unidimensionality.

### Data_3.2.1.2_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to list of some expressions used in study 2 (as stated in the Many Labs 3 protocol) from Many Labs 3 that was checked for reliability and unidimensionality.

### Data_3.7.1_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to argument quality used in study 7 (as stated in the Many Labs 3 protocol) from Many Labs 3 that was checked for reliability and unidimensionality.

### Data_3.7.2_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to need for cognition used in study 7 (as stated in the Many Labs 3 protocol) from Many Labs 3 that was checked for reliability and unidimensionality.

### Data_3.8.2_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to conscientousness used in study 8 (as stated in the Many Labs 3 protocol) from Many Labs 3 that was checked for reliability and unidimensionality.

### Data_5.1.1_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to verbal ability used in study 1 (as ordered in Table 1. in the Many Labs 5 manuscript) from Many Labs 5 that was checked for reliability and unidimensionality.

### Data_5.1.2_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to quantitative ability used in study 1 (as ordered in Table 1. in the Many Labs 5 manuscript) from Many Labs 5 that was checked for reliability and unidimensionality.

### Data_5.4_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to antogonist trait description used in study 4 (as ordered in Table 1. in the Many Labs 5 manuscript) from Many Labs 5 that was checked for reliability and unidimensionality.

### Data_5.7_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to willingness to reconcile used in study 7 (as ordered in Table 1. in the Many Labs 5 manuscript) from Many Labs 5 that was checked for reliability and unidimensionality.

### Data_5.9.1_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to free will used in study 9 (as ordered in Table 1. in the Many Labs 5 manuscript) from Many Labs 5 that was checked for reliability and unidimensionality.

# AnalysisData

### cfa_results.csv

*Explanation:* contains the results obtained from the exploratory unidimensionality check analyses based on factor analyses of the measures related to the data_1.10_clean, data_1.11_clean, data_1.12.3.1_clean, data_1.12.3.2_clean,data_2.12.1_clean, data_2.12.2_clean, data_2.12.3_clean, data_2.15_clean, data_2.20_clean, data_2.23_clean, data_3.2.1.1_clean, data_3.2.1.2_clean, data_3.7.1_clean, data_3.7.2_clean, data_3.8.2_clean, data_5.1.1_clean, data_5.1.2_clean, data_5.7_clean, & data_5.9.1_clean datasets. This is part of the pre-registered exploratory analyses documented in appendix A.

*Variables:*

-   **RMSEA:** the rmsea value for a single factor confirmatory factor analysis model when in turn applied to each of the datasets under consideration.

-   **N factors:** the number of factors selected for each of the datasets based on the parallel test for selecting factors.

-   **Single_factor:** a boolean variable indicating whether or not the measure was unidimensional based on the criteria that either the rmsea for the single factor model was smaller than .08, or the parallel test returned a single factor solution.

### data_H1.csv DELETE???!!!

*Explanation*: Contain all of the data from the coded_data_original_shortened.csv and coded_data_replications.csv except for an empty column that was meant to contain any types of reported psychometric evidence not included in the multiple choice response options, since it caused codebook generation issues. The two datasets of which this dataset was composed were bound so that each row contained both the coded data of the original article and the coded data on the same measure as described in the replication protocols. An additional variable was added which indicated for the multiple item measures only whether or not any form of reliability evidence was reported or not as a boolean variable.

### data_H23.csv

*Explanation*: This dataset list contains information on the calculated Cronbach's alpha and the omega reliability coefficient calculated for the responses to the questionnaire items from several of the Many Labs studies, namely those responses encoded in: Data_1.10_clean.csv, data_1.11_clean.csv, data_1.12.3.1_clean.csv, data_1.12.3.2_clean.csv, data_2.10.1_clean.csv, data_2.12.1_clean.csv, data_2.12.2_clean.csv, data_2.12.3_clean.csv, data_2.15_clean.csv, data_2.20_clean.csv, data_2.23_clean.csv, data_3.2.1.1_clean.csv, data_3.2.1.2_clean.csv, data_3.7.2_clean.csv, data_3.8.2_clean.csv, data_5.1.1_clean.csv, data_5.7_clean.csv, data_5.9.1_clean.csv.

*Variables*:

-   **alpha:** Cronbach's alpha calculated for a measure at a particular lab location within a single study/questionnaire as extracted from the Many Labs datasets.

-   **omega:** McDonald's omega calculated for a measure at a particular lab location within a single study/questionnaire as extracted from the Many Labs datasets.

-   **ASE:** Standard error of Cronbach's alpha at a particular lab location, based on formulas 2 & 3 from @duhachek2004AlphaStandardError.

-   **tau:** the tau estimate fo heterogeneity as obtained from a meta-analysis of the Cronbach's alpha of the same measure across multiple lab locations.

-   **Qep:** the p-value of the Q test for heterogeneity as obtained from a meta-analysis of the Cronbach's alpha of the same measure across multiple lab locations.

-   **pi.lb & pi.ub:** the lower and upper bound of the meta-analytic prediction interval for the meta-analysis of the Cronbach's alpha of the same measure across multiple lab locations.

-   **g:** a short reference to the specific Many Labs replication (and if necessary the particular measure and construct) each alpha was associated with.

-   **reporting_index:** index of the row in the data_h456.csv dataset that the indexed measure belongs to.

-   **replication_success:** boolean variable indicating whether or not the measure's related effect was successfully replicated or not when looking at the significance of the meta-analytic p-value.

### data_H23_avg.csv

*Explanation*: For each measure, the mean of the variables from data_H23.csv, and the related g variable. Additionally, some data comparing the reported and replicated alpha was added.

*Variables*:

-   **coefficient_reported**: If present, the reported alpha reliability coefficient in the original article for the measure in this row.

-   **coefficient_difference**: The difference between the reported alpha coefficient and the average coefficient calculated from the replication data (negative values indicate that the reported coefficient is lower than the calculated average)

-   **siginificance_reported_coefficient**: boolean value indicating whether or not the coeffcient_reported value of a measure is outside the 95% percentile of calculated alpha coefficient score of the same measure.

### data_H456_original.csv

*Explanation*: Contain all of the data from the coded_data_original_shortened.csv and one more variable indicating whether reliability was reported or not.

*Variables*:

-   **reliability_reported**: boolean variable indicating whether any response other than nothing or "Not Reported" was given to the **reliability_type** variable.

### data_H456_replication.csv

*Explanation*: Contain all of the data from the coded_data_replications.csv and one more variable indicating whether reliability was reported or not.

*Variables*:

-   **reliability_reported**: boolean variable indicating whether any response other than nothing or "Not Reported" was given to the **reliability_type** variable.

### item_number_change_data.csv

*Explanation:* Contains the many_labs_version, rep_org, title, and mod_5_rev variablefrom the coded_data_replication.csv dataset of the observations where the mod_5_rev variable was TRUE (meaning that the measure's number of items and/or number of response options was modified from original to replication. Additionally, two variables were added to determine whether for that measure the number of items were decreased from original to replication.

*Variables*:

-   **context:** A description of the change in item numbers from original to replication. NA means that for that measure the number of items was not changed.

-   **direction:** A factor indicating whether the number of items was decreased, increased, or no change (NA) from original to replication.

### plot_23_data_alpha.csv

*Explanation*: Contain all of the data from the data_h23.csv that had a calculated alpha level larger than 0. Additionally, it contains a column of alpha scores averaged across many lab locations for each article. Finally, the replication_success was changed to a boolean variable, returning TRUE in case it was coded as "Yes", and as FALSE otherwise.

### plot_23_data_omega.csv

*Explanation*: Contain all of the data from the data_h3_multiple.csv. Additionally, it contains a column of omega scores averaged across many lab locations for each article. Finally, the replication_success was changed to a boolean variable, returning TRUE in case it was coded as "Yes", and as FALSE otherwise.

### plot_456_data.csv

*Explanation*: Contains four variables for each QMP type for all the replication studies and original articles that were coded on QMPs. The rows are ordered to have the original studies first and then the replications, then they are ordered based on QMP type from Definition to Operationalization to Selection to Quantification to Modification to Total. Within each type the QMPs are ordering matches the order of the studies/articles as in the data_h456_replications.csv and data_h456_original.csv.

*Variables:*

-   **QMP_ratio:** float number between 0 and 1 that indicates the QMP ratio based on the initial coding protocol for that QMP type for that study/article.

-   **QMP_ratio_1:** transformed version of the QMP_ratio column, where QMPs for all original articles have a 1 added to their ratio. This was done for the sake of making the figure possible in ggplot as it is right now.

-   **QMP_type:** factor that indicates the QMP type the QMP ratio belongs to (Definition, Operationalization, Selection, Quantification, or Modification).

-   **QMP_ratio_REV:** float number between 0 and 1 that indicates the QMP ratio based on the revised coding protocol for that QMP type for that study/article.

-   **RepOrg:** factor that indicates whether the QMP ratio belongs to a replication or an original article.

-   **QMP_ratio_1_REV:** transformed version of the QMP_ratio_REV column, where QMPs for all original articles have a 1 added to their ratio. This was done for the sake of making the figure possible in ggplot as it is right now.

### plot_46_data.csv

*Explanation*: Contains four variables for each QMP type (including total) for both replication studies and original articles combined per row. The QMP_ratio column indicates the initial QMP ratio for the specific QMP type for that original article. The QMP_type column indicates the specific QMP type the QMP_ratio column belongs to. The QMP_Rep_ratio column indicates the initial QMP ratio for the specific QMP type for that replication study. The QMP_Rep_type column indicates the specific QMP type the QMP_Rep_ratio column belongs to. The rows are ordered so that within each QMP type it matches the order of the studies/articles as in the data_h456_replications.csv and data_h456_original.csv.

### plot_46_data_REV.csv

*Explanation*: Contains four variables for each QMP type (including total) for both replication studies and original articles combined per row. The QMP_ratio column indicates the revised QMP ratio for the specific QMP type for that original article. The QMP_type column indicates the specific QMP type the QMP_ratio column belongs to. The QMP_Rep_ratio column indicates the revised QMP ratio for the specific QMP type for that replication study. The QMP_Rep_type column indicates the specific QMP type the QMP_Rep_ratio column belongs to. The rows are ordered so that within each QMP type it matches the order of the studies/articles as in the coded_data_replications.csv and coded_data_original_shortened.csv.