### 2025-08-25 Data/Data_package_description.Rmd
adjusted the supplementary materials to reflect the new project structure, fixed references, and removed depricated files.

<!-- Provide here a full description of each dataset in the project not captured by the automated dataset-specific codebooks. -->

# InputData

### Coded_Data_Initial_raw.csv

*Origin*: This dataset was created originally for this research project. It was obtained using the Measurement_Error_Reporting_Coding_Protocol.pdf that can be found in the SupplementaryMaterials/CodingProtocols folder of this project on the Many Labs protocols mentioned in the texts and the related original articles for each of the many labs replications.

*Explanation*: The information in this dataset was used for analyses relating to QMP ratios among original articles and replication studies. It additionally contains peripheral information about the articles and study in it that were used throughout the analyses and for descriptives. Questions related to modification had 'no modification' as an additional answer option, in case that specific aspect of the measure was explicitly not modified, but might possibly have been applicable, because it was a relevant aspect albeit unmodified. But these operated equivalently to the Not applicable response with respect to the calculation of the QMP ratio. They are included here for future reference.

*Variables*:

-   **StartDate, EndData, Status, IPAdress, Progress, Duration..in.seconds., Finished, RecordedDate, Responseid, RecipientLastName, RecipientFirstName, RecipientEmail, ExternalReference, LocationLatitude, LocationLongitude, DistributionChannel,** & **UserLanguage:** automated Qualtrics metadata, not used in study.

-   **Many_labs_version:** Indicates using a number the Many Labs version the study/replication under investigation belongs to.

-   **research_type:** Indicates using text whether a coded measurement comes from a Replication protocol, or an Original article.

-   **research_title:** the title of the replicated study as it appears in the associated Many Labs protocol/original article.

-   **measure_name:** the measure's name as stated within the Many Labs protocol/original article, or a short description based on what was said in the many labs protocol/original article.

-   **variable_name:** the name of the variable the measure was intended to measure based on the description in the many labs protocol/original article.

-   **Q58:** Boolean variable indicating whether or not the variable assessed by the measure was assessed as [part of a larger measure intending to measure multiple variables/dimensions of a construct.

-   **variable_order:** a number indicating the order of appearance for the measure within the total number of measures for that study/article.

-   **N:** the number of people that took the measure as indicated in the many labs report/original article.

-   **N_items:** a three choice option variable indicating whether a measure was single item, multiple item, or not a scale expected to have regular psychometric properties.

-   **hypothesis_support:** question indicating whether or not the Many Labs replications found support for the original hypothesis at a meta-analytic alpha level of .05 in the same direction as the original effect.

-   Reliability Coefficient Variables

    -   **reliability_type:** for multiple item measures only, a question asking if a reliability coefficient was reported, and if so which one with a couple of options and the option to select Other.

    -   **reliability_type_5_TEXT:** for questions that answered Oher on the reliability_type question, a text description of the specific reliability index.

    -   **reliability_coeff:** for all measures with a reliability coefficient, the value of that coefficient as reported in the Many Labs protocol or original article.

-   QMP Variables

    -   **Def1:** A boolean variable with an additional Not Applicable option, indicating whether or not a psychological/sociological definition was given within the protocol study description or original article to the measured variable from the measure.

    -   **Op_version:** if described within the protocol or article, the name and citation for the specific version of the measure that was used.

    -   **Op1:** A boolean variable with an additional Not Applicable option, indicating if the specific version of the measure is described.

    -   **Op2:** A boolean variable with an additional Not Applicable option, indicating whether or not the administration format of the measure and the environment the measure was taken in was described.

    -   **Op3:** A boolean variable with an additional Not Applicable option, indicating whether or not the steps taken in administering the measure were described or not.

    -   **Op4:** A boolean variable with an additional Not Applicable option, indicating whether or not the authors justified the way they conducted the measurement and why the specific choice was necessary to properly measure what they were investigating.

    -   **Op5:** A boolean variable with an additional Not Applicable option, indicating whether or not example items are presented in the text, or in an appendix.

    -   **Sel_existing:** Variable indicating whether existing articles were used, the measure was newly created for the study, or if it was not clearly stated.

    -   **Sel_existing_2_TEXT:** A textual description giving the (general) name or explanation of which previous measure was used, in case an already existing measure was used.

    -   **Sel1:** A boolean variable with an additional Not Applicable option, indicating whether or not the source of the scale is provided, including if it is new that this is clearly stated.

    -   **Sel2:** A boolean variable with an additional Not Applicable option, indicating whether or not reasoning is given for choosing a particular measure either existing or creating a new one.

    -   **Sel3:** A boolean variable with an additional Not Applicable option, indicating whether or not psychometric evidence for the validity of a scale was presented.

    -   **Sel4:** A boolean variable with an additional Not Applicable option, indicating whether or not there is psychometric evidence for the measure outside of the main analyses of the protocol or study, this can be either from an earlier study, or from a pilot study testing the measure.

    -   **Sel_psych_evidence:** A variable meant to indicate whether an analysis of the measure's validity was applicable, and if so whether or not it was there, and if so too, what type of psychometric evidence was reported.

    -   **Sel_psych_evidence_4_TEXT:** A textual description of the specific analyses of validity that was used in case the Other, namely: option was picked in the last question.

    -   **Quant1:** A boolean variable with an additional Not Applicable option, indicating whether or not the number of items in the measure are described.

    -   **Quant2:** A boolean variable with an additional Not Applicable option, indicating whether or not the number of response options for the items in the measure are described.

    -   **Quant3:** A boolean variable with an additional Not Applicable option, indicating whether or not recoding of responses is indicated for multiple item measures.

    -   **Quant4:** A boolean variable with an additional Not Applicable option, indicating whether or not the aggregation of the scores on the items on a multiple item measure into a single index was described.

    -   **Mod_check:** A variable indicating whether or not measurement modifications were mentioned at all, and if so whether the measurement was modified or not.

    -   **Mod1:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect) and Not Applicable option, indicating whether or not any modification to the administration format was mentioned or not.

    -   **Mod2:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect) and Not Applicable option, indicating whether or not there is validity evidence to support the change in the administration format.

    -   **Mod3:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect) and Not Applicable option, indicating whether or not any modification to the language of the questionnaire was mentioned or not.

    -   **Mod4:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect) and Not Applicable option, indicating whether or not there is validity evidence for the translated measure.

    -   **Mod5:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect) and Not Applicable option, indicating whether or not any a change in the number of items or response options deviated from the orginal source.

    -   **Mod6:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect) and Not Applicable option, indicating whether or not there is validity evidence for the measure with a change in the number of items or response options.

    -   **Mod_time:** A variable indicating whether or not the moment the modifications was reported, and if so whether this occurred before or after the data collection for the study.

### Coded_Data_Revised_raw.csv

*Origin*: This dataset was created originally for this research project. It was obtained using the Measurement_Error_Reporting_Revised_Coding_Protocol.pdf that can be found in the SupplementaryMaterials/CodingProtocols folder of this project on the Many Labs protocols mentioned in the texts and the related original articles for each of the many labs replications (the same sample as was used to obtain the Coded_Data_Initial_raw.csv).

*Explanation*: After discussion with the supervisor based on the experiences of the initial coding round it was determined that several of the demarcation for meeting several QMP criteria were likely too strict, and thus several of the more strictly coded questions were revised and coded again using a new coding protocol. The goal being to offer an as lenient as possible overview of the errors in reporting in the investigated literature.

*Variables*:

-   **StartDate, EndData, Status, IPAdress, Progress, Duration..in.seconds., Finished, RecordedDate, Responseid, RecipientLastName, RecipientFirstName, RecipientEmail, ExternalReference, LocationLatitude, LocationLongitude, DistributionChannel,** & **UserLanguage:** automated Qualtrics metadata, not used in study.

-   **Many_labs_version:** Indicates using a number the Many Labs version the study/replication under investigation belongs to.

-   **research_type:** Indicates using text whether a coded measurement comes from a Replication protocol, or an Original article.

-   **research_title:** the title of the replicated study as it appears in the associated Many Labs protocol/original article.

-   **variable_order:** a number indicating the order of appearance for the measure within the total number of measures for that study/article.

-   QMP Variables

    -   **Op1:** A boolean variable with an additional Not Applicable option, indicating if the specific version of the measure is described, or the item in a single item measure is itself described in sufficient detail.

    -   **Op2:** A boolean variable with an additional Not Applicable option, indicating whether or not the administration format of the measure and the environment the measure was taken in was described generally in as far as it is relevant to the construct.

    -   **Op5:** A boolean variable with an additional Not Applicable option, indicating whether or not example items are presented in the text, or in an (online) appendix.

    -   **Sel1:** A boolean variable with an additional Not Applicable option, indicating whether or not the source of the scale is provided. In case the scale is new, the context should make this apparent such that no source needs to be provided for this scale.

    -   **Sel3:** A boolean variable with an additional Not Applicable option, indicating whether or not psychometric evidence for the validity of a scale was presented. Or in case such evidence is not relevant for that particular measure, a statement that clarifies this.

    -   **Sel_psych_evidence:** A variable meant to indicate whether an analysis of the measure's validity was applicable, and if so whether or not it was there, and if so too, what type of psychometric evidence was reported.

    -   **Sel_psych_evidence_4_TEXT:** A textual description of the specific analyses of validity that was used in case the Other, namely: option was picked in the last question.

    -   **Quant1:** A boolean variable with an additional Not Applicable option, indicating whether or not the number of items in the measure are described.

    -   **Quant2:** A boolean variable with an additional Not Applicable option, indicating whether or not the number of response options for the items in the measure are described.

    -   **Quant3:** A boolean variable with an additional Not Applicable option, indicating whether or not recoding of responses is indicated for multiple item measures.

    -   **Mod_check:** A variable indicating whether or not there were known modifications to the measurement before use in the research, and if so whether the measurement was modified or not.

    -   **Mod1:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect) and Not Applicable option, indicating whether or not any modification to the administration format of the overarching study protocol was mentioned or not.

    -   **Mod2:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect, or the modification was of little impact) and Not Applicable option, indicating whether or not there is validity evidence to support the change in the administration format.

    -   **Mod3:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect) and Not Applicable option, indicating whether or not any modification to the language of the overarching study protocol was mentioned or not.

    -   **Mod4:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect, or the modification was of little impact) and Not Applicable option, indicating whether or not there is validity evidence for the translated measure.

    -   **Mod5:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect) and Not Applicable option, indicating whether or not any a change in the number of items or response options deviated from the original source, for those items relevant to the original source.

    -   **Mod6:** A boolean variable with an additional No Modification (in case there was clear that there was no modification to this aspect, or the modification was of little impact) and Not Applicable option, indicating whether or not there is validity evidence for the measure with a change in the number of items or response options.

### Coded_Data_Vignette_raw.csv

*Origin*: This dataset was created originally for this research project. It was obtained using the Vignette_Item_Coding_Protocol.pdf that can be found in the SupplementaryMaterials/CodingProtocols folder of this project on the Many Labs protocols mentioned in the texts and the related original articles for each of the many labs replications (the same sample as was used to obtain the Coded_Data_Initial_raw.csv).

*Explanation*: This dataset was collected after the initial round of coding with the intention to control on certain analyses whether items described a vignette and not an ordinary questionnaire, which might have different standards with regards to measurement reporting and reliability. As of yet, it has remained unused in any analyses

*Variables*:

-   **RecordedDate**: Records the date the variable was coded, was not used further.

-   **Vignette**: Boolean variable indicating whether or not a particular measure was inseparably tied to a vignette describing some form of quoted text, visual, imagined scenario, cognitive task, or game.

### Data_ML1.csv

*Origin*: This dataset was obtained from the Datasets.zip archive as "Full_Dataset_De-Identified.sav", from the following OSF page: <https://osf.io/pqf9r/>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Ratliff, K., Vianello, M., Adams Jr, R. B., Bahník, S., Bernstein, M. J., . . . Nosek, B. A. (2014). Investigating variation in replicability: A "many labs" replication project. Social Psychology, 45 (3), 142--152. <https://doi.org/10.1027/1864-9335/a000178>, the dataset itself has the following doi: 10.17605/OSF.IO/PQF9R.

*Alterations*: the only alterations made to the dataset at this stage were: the name of the dataset was changed, and the format was changed from .sav to .csv.

### Data_2.2.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "Structure & Goal Pursuit (Kay et al., 2014)/Kay.1/Global/Data/" as "Kay_1_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.3.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "Incidental Disfluency (Alter et al., 2007)/Alter.1/Global/Data/" as "Alter_1_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.4.1.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "Moral Foundations (Graham et al., 2009)/Graham.1/Global/Data/" as "Graham_1_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.4.2.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "Moral Foundations (Graham et al., 2009)/Graham.2/Global/Data/" as "Graham_2_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.8.2.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "Disgust & Homophobia (Inbar et al., 2009)/Inbar.1b/Global/Data/" as "Inbar_1b_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.10.1.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "Social Value Orientation (Van Lange et al., 1997)/vanLange.1/Global/Data/" as "vanLange_1_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.12.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "SMS & Well-Being (Anderson et al., 2012)/Anderson.1/Global/Data/" as "Anderson_1_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.15.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "Position & Power (Giessner & Schubert, 2007)/Giessner.1/Global/Data/" as "Giessner_1_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.19.1.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "Actions are Choices (Savani et al., 2010)/Savani.3a/Global/Data/" as "Savani_3a_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.20.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "Intuitive Reasoning (Norenzayan et al., 2002)/Norenzayan.1/Global/Data" as "Norenzayan_1_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_2.23.csv

*Origin*: This dataset was obtained from the OSFdata.zip archive in file path "/Moral Cleansing (Zhong & Liljenquist, 2006)/Zhong.1/Global/Data/" as "Zhong_1_study_global_include_all_CLEAN_CASE.csv", from the following OSF page: <https://osf.io/ag2pd>.

*Attribution*: The datatset is related to the following article: Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams Jr, R. B., Alper, S., . . . Bahnìk, Š. (2018). Many labs 2: Investigating variation in replicability across samples and settings. Advances in Methods and Practices in Psychological Science, 1 (4), 443--490. <https://doi.org/10.1177/2515245918810225>, the dataset itself has the following doi: 10.17605/OSF.IO/FANRE.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_ML3.csv

*Origin*: This dataset was obtained from the ML3 Final Data.zip archive in the base folder as "ML3AllSitesandmTurk.csv", from the following OSF page: <https://osf.io/bxw8j>.

*Attribution*: The datatset is related to the following article: Ebersole, C. R., Atherton, O. E., Belanger, A. L., Skulborstad, H. M., Allen, J. M., Banks, J. B., . . . Nosek, B. A. (2016). Many labs 3: Evaluating participant pool quality across the academic semester via replication. Journal of Experimental Social Psychology, 67, 68--82. <https://doi.org/10.1016/j.jesp.2015.10.012>.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_3.5.csv

*Origin*: This dataset was obtained from the ML3 Final Data.zip archive in the base folder as "ML3StroopData.csv", from the following OSF page: <https://osf.io/bxw8j>.

*Attribution*: The datatset is related to the following article: Ebersole, C. R., Atherton, O. E., Belanger, A. L., Skulborstad, H. M., Allen, J. M., Banks, J. B., . . . Nosek, B. A. (2016). Many labs 3: Evaluating participant pool quality across the academic semester via replication. Journal of Experimental Social Psychology, 67, 68--82. <https://doi.org/10.1016/j.jesp.2015.10.012>.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_5.1.csv

*Origin*: This dataset was obtained as "ML5 Alb 5 Revised Protocol In Lab.xlsx", from the following OSF page: <https://osf.io/64njs>.

*Attribution*: The datatset is related to the following article: Ebersole, C. R., Mathur, M. B., Baranski, E., Bart-Plange, D.-J., Buttrick, N. R., Chartier, C. R., . . . Nosek, B. A. (2020). Many labs 5: Testing pre-data-collection peer review as an intervention to increase replicability. Advances in Methods and Practices in Psychological Science, 3 (3), 309--331. <https://doi.org/10.1177/2515245920958687>.

*Alterations*: the only alterations made to the dataset at this stage were: the name of the dataset was changed, and the format was changed from .xlsx to .csv.

### Data_5.4.csv

*Origin*: This dataset was obtained as "dataset_LoBue_raw.sav", from the following OSF page: <https://osf.io/w7ft9>.

*Attribution*: The datatset is related to the following article: Ebersole, C. R., Mathur, M. B., Baranski, E., Bart-Plange, D.-J., Buttrick, N. R., Chartier, C. R., . . . Nosek, B. A. (2020). Many labs 5: Testing pre-data-collection peer review as an intervention to increase replicability. Advances in Methods and Practices in Psychological Science, 3 (3), 309--331. <https://doi.org/10.1177/2515245920958687>.

*Alterations*: the only alterations made to the dataset at this stage were: the name of the dataset was changed, and the format was changed from .sav to .csv.

### Data_5.5.csv

*Origin*: This dataset was obtained as "Genova_ML5.csv", from the following OSF page: <https://osf.io/prj7y>.

*Attribution*: The datatset is related to the following article: Ebersole, C. R., Mathur, M. B., Baranski, E., Bart-Plange, D.-J., Buttrick, N. R., Chartier, C. R., . . . Nosek, B. A. (2020). Many labs 5: Testing pre-data-collection peer review as an intervention to increase replicability. Advances in Methods and Practices in Psychological Science, 3 (3), 309--331. <https://doi.org/10.1177/2515245920958687>.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_5.7.csv

*Origin*: This dataset was obtained as "ml5_computed_dataset.csv", from the following OSF page: <https://osf.io/da8bh>.

*Attribution*: The datatset is related to the following article: Ebersole, C. R., Mathur, M. B., Baranski, E., Bart-Plange, D.-J., Buttrick, N. R., Chartier, C. R., . . . Nosek, B. A. (2020). Many labs 5: Testing pre-data-collection peer review as an intervention to increase replicability. Advances in Methods and Practices in Psychological Science, 3 (3), 309--331. <https://doi.org/10.1177/2515245920958687>.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

### Data_5.9.csv

*Origin*: This dataset was obtained as "ML5_fulldataset_with_suspicion.csv", from the following OSF page: <https://osf.io/5w6e7>.

*Attribution*: The datatset is related to the following article: Ebersole, C. R., Mathur, M. B., Baranski, E., Bart-Plange, D.-J., Buttrick, N. R., Chartier, C. R., . . . Nosek, B. A. (2020). Many labs 5: Testing pre-data-collection peer review as an intervention to increase replicability. Advances in Methods and Practices in Psychological Science, 3 (3), 309--331. <https://doi.org/10.1177/2515245920958687>.

*Alterations*: the only alteration made to the dataset at this stage was: the name of the dataset was changed.

# IntermediateData

### coded_data_original.csv

*Explanation*: contains all coded data on the original articles with renamed columns, fixed bad data, restructured to match the ordering of the protocols more eligibly, recoded to be close to suitable for analysis. Additionally, he coded data on 3 measures that did not appear in the replication protocols were removed to allow for easier comparisons between replications and original studies.

*Newly Coded Variables*:

-   **\_REV variables:** these relate to the coding obtained for that particular variable using the revised coding protocol instead of the initial coding protocol.

-   **Inseparable material:** the Vignette variable from the Coded_Data_Vignette_raw.csv dataset.

### coded_data_replications.csv

*Explanation*: contains all coded data on the replications as described in the many labs protocols with renamed columns, fixed bad data, restructured to match the ordering of the protocols more ineligibly, recoded to be close to suitable for analysis.

*Newly Coded Variables*:

-   **\_REV variables:** these relate to the coding obtained for that particular variable using the revised coding protocol instead of the initial coding protocol.

-   **Inseparable material:** the Vignette variable from the Coded_Data_Vignette_raw.csv dataset.

### Data_1.3_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to anchoring used in study 3 (as stated in the Many Labs 1 Proposal_V1.1) from Many Labs 1 that was checked for reliability and unidimensionality.

### Data_1.10_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to system justification used in study 10 (as stated in the Many Labs 1 Proposal_V1.1) from Many Labs 1 that was checked for reliability and unidimensionality.

### Data_1.11_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to intentions to engage used in study 11 (as stated in the Many Labs 1 Proposal_V1.1) from Many Labs 1 that was checked for reliability and unidimensionality.

### Data_1.12.3.1_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to explicit attitudes towards art used in study 12 (as stated in the Many Labs 1 Proposal_V1.1) from Many Labs 1 that was checked for reliability and unidimensionality.

### Data_1.12.3.2_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to explicit attitudes towards math used in study 12 (as stated in the Many Labs 1 Proposal_V1.1) from Many Labs 1 that was checked for reliability and unidimensionality.

### Data_2.10.1_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to social value orientation used in study 10 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

### Data_2.12.1_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to satisfaction with life used in study 12 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

### Data_2.12.2_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to positive affect used in study 12 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

### Data_2.12.3_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to negative affect used in study 12 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

### Data_2.15_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to rated power used in study 15 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

### Data_2.20_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to rule or similarity based judgement used in study 20 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

### Data_2.23_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to cleaning product ratings used in study 23 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

### Data_3.2.1.1_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to list of most expressions used in study 2 (as stated in the Many Labs 3 protocol) from Many Labs 3 that was checked for reliability and unidimensionality.

### Data_3.2.1.2_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to list of some expressions used in study 2 (as stated in the Many Labs 3 protocol) from Many Labs 3 that was checked for reliability and unidimensionality.

### Data_3.7.1_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to argument quality used in study 7 (as stated in the Many Labs 3 protocol) from Many Labs 3 that was checked for reliability and unidimensionality.

### Data_3.7.2_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to need for cognition used in study 7 (as stated in the Many Labs 3 protocol) from Many Labs 3 that was checked for reliability and unidimensionality.

### Data_3.8.2_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to conscientousness used in study 8 (as stated in the Many Labs 3 protocol) from Many Labs 3 that was checked for reliability and unidimensionality.

### Data_5.1.1_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to verbal ability used in study 1 (as ordered in Table 1. in the Many Labs 5 manuscript) from Many Labs 5 that was checked for reliability and unidimensionality.

### Data_5.1.2_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to quantitative ability used in study 1 (as ordered in Table 1. in the Many Labs 5 manuscript) from Many Labs 5 that was checked for reliability and unidimensionality.

### Data_5.4_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to antogonist trait description used in study 4 (as ordered in Table 1. in the Many Labs 5 manuscript) from Many Labs 5 that was checked for reliability and unidimensionality.

### Data_5.7_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to willingness to reconcile used in study 7 (as ordered in Table 1. in the Many Labs 5 manuscript) from Many Labs 5 that was checked for reliability and unidimensionality.

### Data_5.9.1_clean.csv

*Explanation*: The first column of this dataset indicates the lab location where the data was collected as a factor. Besides that, this dataset contains all of the columns that could be identified as item scores on the scale relating to free will used in study 9 (as ordered in Table 1. in the Many Labs 5 manuscript) from Many Labs 5 that was checked for reliability and unidimensionality.

# AnalysisData

### calculated_reliability_lab_data.csv

*Explanation*: This dataset list contains information on the calculated Cronbach's alpha and the omega reliability coefficient calculated for the responses to the questionnaire items from several of the Many Labs studies, namely those responses encoded in: Data_1.10_clean.csv, data_1.11_clean.csv, data_1.12.3.1_clean.csv, data_1.12.3.2_clean.csv, data_2.10.1_clean.csv, data_2.12.1_clean.csv, data_2.12.2_clean.csv, data_2.12.3_clean.csv, data_2.15_clean.csv, data_2.20_clean.csv, data_2.23_clean.csv, data_3.2.1.1_clean.csv, data_3.2.1.2_clean.csv, data_3.7.2_clean.csv, data_3.8.2_clean.csv, data_5.1.1_clean.csv, data_5.7_clean.csv, data_5.9.1_clean.csv.

*Variables*:

-   **alpha:** Cronbach's alpha calculated for a measure at a particular lab location within a single study/questionnaire as extracted from the Many Labs datasets.

-   **omega.tot:** McDonald's omega for the overarching factor as an estimate of the total reliability of a test calculated for a measure at a particular lab location within a single study/questionnaire as extracted from the Many Labs datasets.

-   **omega.hier:** McDonald's hierarchical omega as the estimate of the general factor saturation of a test (not refered to much in study) calculated for a measure at a particular lab location within a single study/questionnaire as extracted from the Many Labs datasets.

-   **ASE:** Standard error of Cronbach's alpha at a particular lab location, based on formulas 2 & 3 from @duhachekAlphasStandardError2004a.

-   **g:** a short reference to the specific Many Labs replication (and if necessary the particular measure and construct) each alpha was associated with.

-   **coded_data_index:** index of the row in the coded_data_original.csv or coded_data_replication.csv dataset that the indexed measure belongs to.

-   **replication_success:** boolean variable indicating whether or not the measure's related effect was successfully replicated or not when looking at the significance of the meta-analytic p-value.

### FA_Albarracin_2008_math.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 1 (as stated by order in Table 1. in the Many Labs 5 manuscript) from Many Labs 5 that was checked for reliability and unidimensionality.

*Variables*:

-   **RMSEA:** the RMSEA of a single factor model fitted to the data of this particular lab.

-   **RMSEA_se:** the standard error of the RMSEA of a single factor model fitted to the data of this particular lab.

-   **N_factors:** the number of factors suggested by the parallel analysis on the data of this particular lab.

-   **unidimensional:** the result of our unidimensionality decision rule (RMSEA \<.08 or N_Factors = 1) for this particular lab.

-   **Tau_chisq_diff**: the chi square difference between the Tau restricted single factor model and the non-tau-restricted single factor model each fitted to the data of this particular lab.

-   **Tau_Df_diff**: the degrees of freedom difference between the Tau restricted single factor model and the non-tau-restricted single factor model each fitted to the data of this particular lab.

-   **Tau_p_diff**: the p value of the chi square test between the Tau restricted single factor model and the non-tau-restricted single factor model each fitted to the data of this particular lab.

-   **mean_spec:** the mean of the fully standardized estimated coefficient change for all the five highest mod index freed parameters for the unidimensional model for this particular lab.

-   **err_chisq_diff**: the chi square difference between the single factor model with the five (sequentially) highest mod index freed parameters freed and the non-freed parameters single factor model each fitted to the data of this particular lab.

-   **err_Df_diff**: the degrees of freedom difference between the single factor model with the five (sequentially) highest mod index freed parameters freed and the non-freed parameters single factor model each fitted to the data of this particular lab.

-   **err_p_diff**: the p value of the chi square test between the single factor model with the five (sequentially) highest mod index freed parameters freed and the non-freed parameters single factor model each fitted to the data of this particular lab.

### FA_Albarracin_2008_verb.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 1 (as ordered in Table 1. in the Many Labs 5 manuscript) from Many Labs 5 that was checked for reliability and unidimensionality.

*Variables*:

-   Same as FA_Albarracin_2008_math.csv

### FA_Anderson_2012_NA.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 12 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

*Variables*:

-   Same as FA_Albarracin_2008_math.csv

### FA_Anderson_2012_PA.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 12 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

*Variables*:

-   Same as FA_Albarracin_2008_math.csv

### FA_Anderson_2012_SWL.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 12 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

*Variables*:

-   Same as FA_Albarracin_2008_math.csv

### FA_Cacioppo_1983_arg.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 7 (as stated in the Many Labs 3 protocol) from Many Labs 3 that was checked for reliability and unidimensionality.

*Variables*:

-   Same as FA_Albarracin_2008_math.csv

### FA_Cacioppo_1983_nfc.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 7 (as stated in the Many Labs 3 protocol) from Many Labs 3 that was checked for reliability and unidimensionality.

*Variables*:

-   Same as FA_Albarracin_2008_math.csv

### FA_Caruso_2012.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 10 (as stated in the Many Labs 1 Proposal_V1.1) from Many Labs 1 that was checked for reliability and unidimensionality.

*Variables*:

-   Same as FA_Albarracin_2008_math.csv

### FA_De_Fruyt_2000.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 8 (as stated in the Many Labs 3 protocol) from Many Labs 3 that was checked for reliability and unidimensionality.

*Variables*:

-   Same as FA_Albarracin_2008_math.csv

### FA_Giessner_2007.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 15 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

*Variables*:

-   Same as FA_Albarracin_2008_math.csv

### FA_Husnu_2010.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 11 (as stated in the Many Labs 1 Proposal_V1.1) from Many Labs 1 that was checked for reliability and unidimensionality.

*Variables*:

-   Same as FA_Albarracin_2008_math.csv

### FA_Monin_2001_most.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 2 (as stated in the Many Labs 3 protocol) from Many Labs 3 that was checked for reliability and unidimensionality.

*Variables*:

-   Same as FA_Albarracin_2008_math.csv

### FA_Monin_2001_some.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 2 (as stated in the Many Labs 3 protocol) from Many Labs 3 that was checked for reliability and unidimensionality.

*Variables*:

-   Same as FA_Albarracin_2008_math.csv

### FA_Norenzayan_2002.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 20 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

*Variables*:

-   Same as FA_Albarracin_2008_math.csv

### FA_Nosek_2002_Art.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 12 (as stated in the Many Labs 1 Proposal_V1.1) from Many Labs 1 that was checked for reliability and unidimensionality.

*Variables*:

-   Same as FA_Albarracin_2008_math.csv

### FA_Nosek_2002_Math.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 12 (as stated in the Many Labs 1 Proposal_V1.1) from Many Labs 1 that was checked for reliability and unidimensionality.

*Variables*:

-   Same as FA_Albarracin_2008_math.csv

### FA_Shnabel_2008.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 7 (as ordered in Table 1. in the Many Labs 5 manuscript) from Many Labs 5 that was checked for reliability and unidimensionality.

*Variables*:

-   Same as FA_Albarracin_2008_math.csv

### FA_Vohs_2008.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 9 (as ordered in Table 1. in the Many Labs 5 manuscript) from Many Labs 5 that was checked for reliability and unidimensionality.

*Variables*:

-   Same as FA_Albarracin_2008_math.csv

### FA_Zhong_2006.csv

*Explanation*: This dataset describes for each lab location of study several unidimensionality, factor analysis, and Cronbach's Alpha assumptions tests for study 23 (as stated in the Many Labs 2 protocol) from Many Labs 2 that was checked for reliability and unidimensionality.

*Variables*:

-   Same as FA_Albarracin_2008_math.csv

### measure_reliability_data.csv

*Explanation*: For each measure, the mean of the variables from calculated_reliability_lab_data.csv, and the related g variable. Additionally, some data comparing the reported and replicated alpha was added.

*Variables*:

-   **alpha:** Cronbach's alpha calculated for a measure across lab locations as extracted from the Many Labs datasets.

-   **omega.tot:** McDonald's omega for the overarching factor as an estimate of the total reliability of a test calculated for a measure across lab locations as extracted from the Many Labs datasets.

-   **omega.hier:** McDonald's hierarchical omega as the estimate of the general factor saturation of a test (not refered to much in study) calculated for a measure across lab locations as extracted from the Many Labs datasets.

-   **ASE:** Standard error of Cronbach's alpha across lab locations, based on formulas 2 & 3 from @duhachekAlphasStandardError2004a.

-   **tau:** the tau estimate fo heterogeneity as obtained from a meta-analysis of the Cronbach's alpha of the same measure across multiple lab locations.

-   **Qep:** the p-value of the Q test for heterogeneity as obtained from a meta-analysis of the Cronbach's alpha of the same measure across multiple lab locations.

-   **pi.lb & pi.ub:** the lower and upper bound of the meta-analytic prediction interval for the meta-analysis of the Cronbach's alpha of the same measure across multiple lab locations.

-   **g:** a short reference to the specific Many Labs replication (and if necessary the particular measure and construct) each alpha was associated with.

-   **coded_data_index:** index of the row in the coded_data_original.csv or coded_data_replication.csv dataset that the indexed measure belongs to.

-   **replication_success:** boolean variable indicating whether or not the measure's related effect was successfully replicated or not when looking at the significance of the meta-analytic p-value.

-   **reported_coefficient**: If present, the reported alpha reliability coefficient in the original article for the measure in this row.

-   **coefficient_difference**: The difference between the reported alpha coefficient and the average coefficient calculated from the replication data (negative values indicate that the reported coefficient is lower than the calculated average)

-   **significance_reported_coefficient**: boolean value indicating whether or not the reported_coeffcient value of a measure is outside the 95% percentile of calculated alpha coefficient score of the same measure.

### QMP_ratio_data.csv

Percentage of measures that have a QMP for each of our QMP items and the number of measures for which they are applicable by both original and replication. The data is organized by QMP item in rows, and the percentages & number of applicable measures being the variables for both original and replication. We also calculated a Phi Coefficient as a correlation between the original QMP and replication QMP being present.

*Variables*:

-   **Practice:** a description of the reporting practice for which we count the percentage of QMPs.

-   **QMP_percentage_original:** the QMP percentage for a particular measurement information in the original studies. This is done by dividing the number of measures for which the relevant measurement information was not available by the total number of measures for which the particular item was applicable.

-   **N_applicable_original:** the number of measures in the original studies for which the measurement item is applicable (factor analytic evidence reporting for single-item measures for instance would not be applicable).

-   **QMP_percentage_replication:** the QMP percentage for a particular measurement information in the replication projects. This is done by dividing the number of measures for which the relevant measurement information was not available by the total number of measures for which the particular item was applicable.

-   **N_applicable_replication:** the number of measures in the replication projects for which the measurement item is applicable (factor analytic evidence reporting for single-item measures for instance would not be applicable).

-   **Phi:** the phi coefficient is a correlation between the original QMP percentage and replication QMP percentage.