### 2025-08-25 SupplementaryMaterials/SupplementaryAnalysesScripts/Supplementary_exploratory_version_pre-reg_analyses.Rmd

<!-- Supplementary Analyses A -->

```{r additional setup}
### NOTE: also run the setup code in the manuscript
library(lmerTest)
library(betareg)
library(statmod)
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(17042023)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r data_prep}
# combine replication and original data 
is_replication <- c(rep(FALSE, 77), rep(TRUE, 77))

combi_org_rep <- rbind(coded_data_original, coded_data_replications)

combi_org_rep <- cbind(combi_org_rep, is_replication)

combi_org_rep$has_reliability <- combi_org_rep$reliability_type != "Not Reported" & combi_org_rep$reliability_type != ""

# function to calculate QMP ratios
calculating <- function(Data){
  QMP_data <- data.frame(def_ratio = rep(NA, 154),
                         op_ratio = rep(NA, 154),
                         sel_ratio = rep(NA, 154),
                         quant_ratio = rep(NA, 154),
                         mod_ratio = rep(NA, 154),
                         QMP_ratio = rep(NA, 154),
                         op_REV_ratio = rep(NA, 154),
                         sel_REV_ratio = rep(NA, 154),
                         quant_REV_ratio = rep(NA, 154),
                         mod_REV_ratio = rep(NA, 154),
                         QMP_REV_ratio = rep(NA, 154))

  def_count <- as.numeric(Data$def_1)
  def_count[is.na(def_count)] <- 0
  def_total <- as.numeric(!is.na(Data$def_1))
  QMP_data$def_ratio <- 1 - round(def_count / def_total, 4) 
  QMP_data$def_ratio[is.na(Data$def_ratio)] <- 0

  op_count <- rowSums(Data[16:20], na.rm = TRUE)
  op_total <- rowSums(!is.na(Data[16:20]))
  QMP_data$op_ratio <- 1 - round(op_count / op_total, 4)

  sel_count <- rowSums(Data[23:26], na.rm = TRUE)
  sel_total <- rowSums(!is.na(Data[23:26]))
  QMP_data$sel_ratio <- 1 - round(sel_count / sel_total, 4)

  quant_count <- rowSums(Data[29:32], na.rm = TRUE)
  quant_total <- rowSums(!is.na(Data[29:32]))
  QMP_data$quant_ratio <- 1 - round(quant_count / quant_total, 4)

  mod_count <- rowSums(Data[34:39], na.rm = TRUE)
  mod_total <- rowSums(!is.na(Data[34:39]))
  QMP_data$mod_ratio <- 1 - round(mod_count / mod_total, 4)

  MP_total <- def_total + op_total + sel_total + quant_total + mod_total
  MP_count <- def_count + op_count + sel_count + quant_count + mod_count
  QMP_data$QMP_ratio <- 1 - round(MP_count / MP_total, 4)

  op_REV_count <- rowSums(Data[c(18,19,41,42,43)], na.rm = TRUE)
  op_REV_total <- rowSums(!is.na(Data[c(18,19,41,42,43)]))
  QMP_data$op_REV_ratio <- 1 - round(op_REV_count / op_REV_total, 4)

  sel_REV_count <- rowSums(Data[c(24,26,44,45)], na.rm = TRUE)
  sel_REV_total <- rowSums(!is.na(Data[c(24,26,44,45)]))
  QMP_data$sel_REV_ratio <- 1 - round(sel_REV_count / sel_REV_total, 4)

  quant_REV_count <- rowSums(Data[c(32,48,49,50)], na.rm = TRUE)
  quant_REV_total <- rowSums(!is.na(Data[c(32,48,49,50)]))
  QMP_data$quant_REV_ratio <- 1 - round(quant_REV_count / quant_REV_total, 4)

  mod_REV_count <- rowSums(Data[52:57], na.rm = TRUE)
  mod_REV_total <- rowSums(!is.na(Data[52:57]))
  QMP_data$mod_REV_ratio <- 1 - round(mod_REV_count / mod_REV_total, 4)

  MP_REV_count <- def_count + op_REV_count + sel_REV_count + quant_REV_count + mod_REV_count
  MP_REV_total <- def_total + op_REV_total + sel_REV_total + quant_REV_total + mod_REV_total
  QMP_data$QMP_REV_ratio <- 1 - round(MP_REV_count / MP_REV_total, 4)

  return(QMP_data)
}

# calculating the QMP ratios for each QMP item.
QMP_Ratio_data <- calculating(combi_org_rep)

QMP_Ratio_data <- cbind(QMP_Ratio_data, is_replication)

```

# Supplementary Analyses A: Pre-registration & Exploratory Analyses
This appendix contains two things. First of all the pre-registered exploratory analyses. Additionally, it also includes the revised analyses shared in the pre-print for this article as found in: <https://osf.io/2gau8_v1>.

Every hypothesis test used an alpha significance cutoff of .05. No correction for multiple testing was applied, to avoid inflating type II error rate. The generally small sample size and consequently low power of these analyses meant that a smaller alpha would increase the number of false negatives. On top of that, these analyses were considered as exploratory. In this exploratory context, the rate of discovery of any potential true effects was given more value relative to decreasing any false positives compared to typical confirmatory research. That as a consequence obtaining false positives was more likely has been accepted, as false negatives were deemed more harmful to this study's aims.

Some of these analyses were also conducted after removing the potentially confounding factor of measures for which the number of items were reduced in the replications. These results can be found in [Supplementary Analyses C](Supplementary_item_reduced_measure_analyses.Rmd).

# Initial Analysis Goals & Process and Reasoning of Omission
We initially had planned to explore relations between both measurement reporting and reliability with replication outcome. This involved regression models wherein an outcome - either the replication outcome from the Many Labs replications or our QMP indicator - and a predictor - reliability, our QMP indicator, or whether it is an original study or a replication. The idea of doing this was based on earlier research that has shown that reliability is related to replication success. The reason for this is because lower reliability introduces more noise into our data, which leads to less stable estimates across multiple studies of the same effects, which leads to less replicable effects. We also wanted to check for measurement-based predictors for replicability more broadly, which includes QMPs from the measurement reporting side of things. The reason for doing this was because knowing what predicts replicability can help us understand the various forces that can inhibit replicability and inform us in when and how to choose what to replicate. This last point is something we still explore in the article itself, albeit using descriptive information and more logic-based inferences rather than statistical testing.

The reason we omitted these analyses is largely for two reasons: (1) if we want to model a relationship between two variables, we also need to account for all potentially relevant confounders to get a good estimate; (2) predicting replicability as a goal may not really be all that interesting theoretically or practically. To elaborate on point 1, what we lack in our case is a theoretical model to understand all the various influences on replicability. Without knowing these, we could technically still use reliability as a predictor, but our goal and motivation to check reliability was to test the theoretical relation between reliability and replicability in practice. The "in practice" part becomes difficult however because practice is messy and we have to account for that messiness, but we can't, because we don't know what to account for. Trying to create a model for this or do simulations to figure this out drifts from the main goal of the article: to understand the reporting, reliability, and validity of measurement in original; and addressing confounds to replication success is just simply beyond the article's scope. To elaborate on point 2, while there are some reasons to predict replicability for practical reasons - mostly for deciding which study to replicate. But as we stated in point 1, it is difficult to effectively predict replicability in our study. So then we are left with the potential theoretical interest of predicting replicability. The idea here is that if we want to understand why something may not replicate, then trying to find predictors for replicability should give us an understanding of what makes something replicable or not. But understanding what makes something replicable is missing the goal of why we care for replicability, which to some extent is that we don't care about replicability. What we care about is getting accurate estimates of an effect of interest, or less statistically is we want to have some certainty that our understanding of how much a psychological phenomenon exists - if it exists at all. But if we just look at trying to see how we can get the most replicable effects possible that's easy, we just conduct Stroop-tasks and other basic psychology 101 tests and we can replicate around the clock. Making studies replicable should not be a goal in and of itself, what we desire is to get accurate estimates of a phenomenon of interest.

There is another key element from the first pre-print version of the article and the pre-registration: we developed a QMP index to allow us to test the relation between measurement reporting and replication outcome statistically. The QMP index was essentially the same as the QMP ratio presented in the article, the number of QMP items that were rated as insufficiently reported on divided by the total number of applicable items. There were two differences, one is that we also created QMP ratios per specific QMP category (definition, selection/creation, operationalization, quantification, and modification); second is that we created this QMP ratio as a variable, thus treating our QMP items collectively as an index. Having an index means that it should be validated in some way (a point we make throughout the article too). Now, we did not make a measure of a construct, so construct validity was not a concern of ours. However, an indicator should have some predictive validity, which requires us to validate that that is the case before we actually use the measure. We did not have the means nor the method nor any idea on how to do this within our study in a way that would be appropriate. We can propose that QMPs should predict replciabiltiy and then show that this is the case, but if our analysis itself is to investigate whether this is the case, then we can't treat it as validating our measure, because before doing that analysis there was no guarantee that it should be related, and thus we can't use observing that relationship as evidence that the measure worked as intended.

So, what would we advice future researchers interested in probing relations with replication and/or creating a QMP index. First, we would advise you to only probe relations with replication if you have a practical reason for wanting to predict replicability. If you want to look into this relation for theoretical reasons, we would advise you to switch your goals. Instead, try and focus on understanding what makes for accurate and robust estimates. This is closer to what we are actually interested in, so if there is something we should want to understand on a deeper level, it is this. Doing so would require taking a more meta-analytic approach to your outcomes, as that can give you more undeerstanding of an effect's robustness compared to the overall p-value that we used as an indicator of replication success. Second, if you want to create a predictve index or even just a general useful descriptive rating scale based on reporting practices, such as QMPs, we would advise you to first of all make this its own study (or studies) as this is likely an involved process. Then, you may want to look at measurement quality standards, such as the @americaneducationalresearchassociationStandardsEducationalPsychological2014 guidelines, to get something more substantially theoretical as a background for what you mean with what should be reported about a measure. Finally, make use of existing research on measurement reporting such as this one (you can look at our coding protocol supplements and the revisions we did to understand what problems may be smart to avoid), but also @flakeConstructValidationSocial2017, @maireadshawMeasurementPracticesLargescale2020, and @flakeConstructValidityValidity2022 have done work on this and have openly available materials. While we made use of these materials, in our goal to make a QMP indicator, we regret mostly (1) not deciding on a strong definition of what good measurement reporting should in the end do, as well as (2) an outcome that reflects the impact of good measurement reporting as we defined it.. We therefore decided to leave creating an index behind for this project, but hopefully this information ahs been useful to inform future research.

Below are the results for our pre-registered analyses for those interested. Take all of them with a big grain of salt.

## Hypothesis 1
As a first step it is worth investigating what the state of measurement error is in original and replication research in psychology. Thus, this study investigated the following:

-   RQ1a. What is the degree of score reliability in replications of psychological research?

-   RQ1b. What is the degree of score reliability in original psychological research?

Since direct replications intend to investigate effects in nearly identical scenarios with nearly identical measures as the original study, it is expected that measurement errors are as present in replication research as in original psychological research.

-   H1. No difference is present in the degree of score reliability between replication research and original psychological research.

### Analysis
There were too few reported reliability coefficients, therefore the test for hypothesis 1 was changed from the pre-registration, and we won't try to do that pre-reg analysis because they are simply to few. Thus for testing hypothesis 1, instead of comparing the reported Cronbach's alpha coefficient between original and replication, the total number of reported reliability coefficients were compared instead. This total also included reliability coefficients other than Cronbach's alpha. This comparison was implemented via a chi-square test using the base R stats package [@R-base].

```{r Hypothesis1Model, include = FALSE}
# New H1 test
# A chi square test for difference in N reliability reported for multiple 
# item measures.
H1_test_result <- chisq.test(combi_org_rep$is_replication[combi_org_rep$N_items == "multiple item measure"], combi_org_rep$has_reliability[combi_org_rep$N_items == "multiple item measure"])

```

Reliabilities were found to be more commonly reported in original articles compared to replication protocols ($\chi^{2}$(`r apa_num(H1_test_result$parameter)`) = `r apa_num(H1_test_result$statistic)`, *p* = `r apa_p(H1_test_result$p.value)`). These results are contradictory to hypothesis 1.

## Hypothesis 2
While reliability estimates are generally considered a measure specific feature, this is false. They are context dependent, and as a result the estimates of an effect can differ across contexts [@cho2015cronbach; @pauly2018resampling]. With large enough differences across contexts, the effect estimates between different replication attempts could deviate from each other beyond direct comparison. To investigate this potential problem, the variation in score reliability is assessed for each measure when used in different contexts.

-   RQ2. To what degree do reliability estimates differ within a replication set of the same original study?

Even though reliability estimates are context dependent, sufficient protocol structure can counteract the variance in reliability estimates. It is possible to reduce variability as the result of context using structured protocols, as was done in the Many Labs projects this study analyzed.

-   H2. There is no significant variation in the reliability estimates of replications of the same original study.

### Analysis
Using the lmer function from the lme4 package [@R-lme4], a multilevel null model with the Cronbach's alpha or omega coefficient of a replication as dependent was specified, where the random intercept was grouped by replication set. The formulas for this model, where $alpha_{ij}$ represents the calculated Cronbach's alpha coefficient of a single study at one lab location with $i = 1,..., n_{locations}$ and $j = 1,...,n_{replications}$ look as follows:

```{=tex}
\begin{align}
& Level 1: alpha_{ij} = \beta_{0j} + e_{ij} \qquad\qquad\quad e_{ij} \sim N(0, \sigma_{e}^{2}) (\#eq:level1H2) \\
& Level 2: \beta_{0j} = \gamma_{00} + u_{0j} \qquad\qquad\qquad u_{0j} \sim N(0, \sigma_{u0}^{2}) (\#eq:level2H2) \\
& Mixed: alpha_{ij} = \gamma_{00} + u_{0j} + e_{ij} (\#eq:mixedH2)
\end{align}
```

The purpose of this analysis was to investigate variability in calculated reliability coefficients within and between replication sets.

```{r Hypothesis2Model, include = FALSE}
# intercept only model to check the within and between group variance.
H2_test_result_alpha <- lmer(formula = alpha ~ 1 + (1|g), data = calculated_reliability_lab_data)

# and same analyses for omega
# intercept only model to check the within and between group variance.
H2_test_result_omega <- lmer(formula = omega.tot ~ 1 + (1|g), data = calculated_reliability_lab_data[complete.cases(calculated_reliability_lab_data),])
```

```{r ICCCalculations, include = FALSE}
# alpha
within_group_variance_alpha <- attr(VarCorr(H2_test_result_alpha), "sc")^2
between_group_variance_alpha <- attr(VarCorr(H2_test_result_alpha)$g, "stddev")^2

ICC_alpha <- between_group_variance_alpha / (within_group_variance_alpha + between_group_variance_alpha)

# omega
within_group_variance_omega <- attr(VarCorr(H2_test_result_omega), "sc")^2
between_group_variance_omega <- attr(VarCorr(H2_test_result_omega)$g, "stddev")^2

ICC_omega <- between_group_variance_omega / (within_group_variance_omega + between_group_variance_omega)
```

The between-group and within-group variance for Cronbach's alpha were approximately `r apa_num(between_group_variance_alpha, digits = 3)` and `r apa_num(within_group_variance_alpha, digits = 3)` respectively. The resulting ICC, obtained using formula \@ref(eq:ICC), was approximately `r apa_num(ICC_alpha, digits = 3)`. This suggests that variance was larger between replication sets than within. The same analysis for omega coefficients even showed a relatively larger degree of between group variance (between-group variance `r apa_num(within_group_variance_omega, digits = 3)`, within-group variance `r apa_num(between_group_variance_omega, digits = 3)`, ICC = `r apa_num(ICC_omega, digits = 3)`). Both results are in line with hypothesis 2, although caution in interpreting this result is advised due to the small sample size and the lack of an available inferential test.

## Hypothesis 3
This study investigated the relationship between reliability and replicability, by using data from real life replication projects.

-   RQ3. What is the association between replication study score reliability and replication outcome?

Greater score-reliability means greater consistency in estimation as variance around the estimate of the true effect is decreased. Consequently measures across different occasions, including between original and replication, should also become more consistent [@nunnaly1994assessment]. When original and replication research share consistent result, and as long as the studied effect is true, then on average their statistical conclusions should converge.

-   H3. Score reliability in replication research is positively associated with successful replication of an original finding.

### Analysis
In order to test hypothesis 3, the average calculated Cronbach's alpha level within a replication set is used as a predictor of replication success. Replication success was either set to 1 if it reached meta analytic significance at an alpha of .05 in the same direction as that found in the original study, or 0 otherwise. This test aligns with a logistic regression model. With $i = 1,..., n_{replications}$, $p_{replicated}$ being binomially distributed, and $alpha$ referring to the averaged calculated Cronbach's alpha coefficient, the following regression equation was used to model the relation:

```{=tex}
\begin{equation}
f(replicated_{i}) = \beta_{0} + \beta_{1} alpha_{i} (\#eq:H3Main)
\end{equation}
```
Where:

```{=tex}
\begin{equation}
f(.) = log\frac{p_{replicated}}{1 - p_{replicated}} (\#eq:LogitLink)
\end{equation}
```

```{r Hypothesis3MainTest, include = FALSE}
## The main model to test hypothesis 3, through a logistic regression model.
H3_test_result_main_alpha <- apa_print(glm(formula = replication_success ~ 1 + alpha, 
                                      family = binomial(), data = measure_reliability_data))

# and for omega
H3_test_result_main_omega <- apa_print(glm(formula = replication_success ~ 1 + 
                                     omega.tot, family = binomial(), 
                                     data = measure_reliability_data))

# alpha says negative, omega says positive, neither are significant though.

```

In this data, replication attempts at different lab locations can be seen as nested within a replication set. In order to see if this nesting might impact the analyses of hypothesis 3, multilevel logistic regression models were included as sensitivity checks. The dependent variable remains replication success at the replication set level, while Cronbach's Alpha is no longer averaged across locations. The models used were a multilevel logistic random intercept and random slope model.

```{r Hypothesis3MultilevelModelSensitivityAnalyses, include = FALSE}
## sensitivity analyses.
# running a random intercept multilevel logistic model
H3_test_result_int_sensitivity_alpha <- apa_print(glmer(formula = replication_success ~ 1 
                                                + alpha + (1|g), family = 
                                                binomial(), data = 
                                                measure_reliability_data))

# and for omega
H3_test_result_int_sensitivity_omega <- apa_print(glmer(formula = replication_success ~ 1 
                                                + omega.tot + (1|g), family = 
                                                  binomial(), data = 
                                                  measure_reliability_data))

# same as for non-multilevel model

# running a random slope multilevel logistic model
H3_test_result_slp_sensitivity_alpha <- apa_print(glmer(formula = replication_success ~ 1 
                                                + alpha + (1 + alpha|g), 
                                                family = binomial(), data =
                                                measure_reliability_data))

# and for omega
H3_test_result_slp_sensitivity_omega <- apa_print(glmer(formula = replication_success ~ 1 
                                                + omega.tot + (1 + omega.tot|g), 
                                                family = binomial(), data =
                                                measure_reliability_data))

# alpha and omega did not converge.
```

The results of the logistic and multilevel logistic models are shown in Table \@ref(tab:ReplicabilityReliabilityTestTable) for both Cronbach's alpha and omega reliability coefficients.

```{r ReplicabilityReliabilityTestTable, warning = FALSE}
# creating the dtaframe for the table
H3_table_dataframe <- rbind(H3_test_result_main_alpha$table, H3_test_result_int_sensitivity_alpha$table, H3_test_result_main_omega$table, H3_test_result_int_sensitivity_omega$table)

# adding rows to give indication of specific sections (apa_table being awkward)
H3_table_for_printing <- rbind(c("Alpha", "", "", "", ""), c("Logistic Regression Model", "", "", "", ""), H3_table_dataframe[1:2,], c("Random-Intercept Multilevel Logistic Model", "", "", "", ""), H3_table_dataframe[(3:4),], c("Omega", "", "", "", ""), c("Logistic Regression Model", "", "", "", ""), H3_table_dataframe[(5:6),], c("Random-Intercept Multilevel Logistic Model", "", "", "", ""), H3_table_dataframe[(7:8),])

# print the table in apa formatting
apa_table(
  H3_table_for_printing, align = c("l", "r", "r", "r", "r")
  , caption = "Model Results from Tests of Hypothesis 3"
  , note = "The Alpha and Omega sections show the models where the independent variable was the calculated Cronbach's alpha and omega coefficient respectively. The random slope model for Cronbach's alpha did not converge, thus it results are not shown here. $^* p < .05$"
  , escape = FALSE, placement= "htp", midrules = c(1, 2, 5, 7, 8, 9, 12))

```

Cronbach's alpha and omega did not significantly relate to replication success in the main logistic regression model (Cronbachs alpha: `r H3_test_result_main_alpha$full_result$alpha`; omega: `r H3_test_result_main_omega$full_result$omega`), nor in any of the multilevel sensitivity analyses models (see Table \@ref(tab:ReplicabilityReliabilityTestTable)). This is contradictory to the expectation in hypothesis 3. However, the results from these analyses should be evaluated with caution, given the small sample number of measures used to estimate these models.

## Hypothesis 4, 5, & 6
Hypothesis 4, 5, & 6 were all tested using a model with the same dependent variable: total QMP ratio. In order to model a ratio dependent variable, beta regression was used. Beta regression models operate similarly to standard generalized linear models with similar interpretations of the results. What differentiates them is that they are suitable to model dependent variables with values in the interval of $(0, 1)$, meaning ratios. Furthermore, they are well suited for modelling heteroskedastic and asymmetrically distributed dependent variables [@cribari2010beta]. The beta regression model was implemented using the betareg function within the betareg package [@R-betareg_a; @R-betareg_b]. Defaults were used for all function parameters besides the model formula and data. All beta regression models used the beta link function, which is as follows:

```{=tex}
\begin{equation}
g(.) = (0, 1) \mapsto \mathbb{R} (\#eq:BetaLink)
\end{equation}
```

All of these models assume that the QMPs are the same across each replication set. The reason being that only one protocol was analysed per replication set, and several of the questions regarding QMPs focus solely on the reporting of relevant measurement information. This assumption was deemed likely, since each replication within a set was based on the same protocol. Furthermore, the goal of these protocols is to ensure that the measurement practices are equal across the different replication studies of the same original effect.

## Hypothesis 4
This study seeks to conceptually replicate some of @flake2022construct findings within the Many Labs replication projects.

-   RQ4a. To what degree are QMPs present in replications of psychological research?
-   RQ4b. To what degree are QMPs present in original psychological research?

The structured and preregistered protocols used for the Many Labs projects were hypothesized to lead to consistent reporting of important measurement details, and as a result contain less QMPs.

-   H4. QMPs are expected to be more frequent in original psychological research than in replication research.

### Analysis
To test hypothesis 4, the total QMP ratios between replication and original research was compared. The model formula for this test looked as follows:

```{=tex}
\begin{equation}
g(QMP\_ratio_{i}) = \beta_{0} + \beta_{1} replication_{i} (\#eq:H4Main)
\end{equation}
```

```{r Hypothesis4Model, include = FALSE}
## Hypothesis 4 beta regression model
# running the model with none revised QMP ratio's
H4_test_results <- betareg(QMP_ratio ~ is_replication, data = QMP_Ratio_data)

# running the models with revised QMP ratio's
H4_test_results_REV  <- betareg(QMP_REV_ratio ~ is_replication, data = QMP_Ratio_data)

```

Where $replication_{i}$ is a dummy coded variable, which was coded as 1 in case the reported QMP ratio belonged to a replications, and 0 if it belonged to an original study.

Results showed that original articles contained a significantly larger proportion of QMPs than replication protocols for the measures of the same effects (beta(`r H4_test_results_REV$df.null`) = `r apa_num(H4_test_results_REV$coefficients$mu[[2]])`, *p* = `r apa_p(summary(H4_test_results_REV)[1]$coefficients$mu[2,4])`), and results based on the initial QMPs showed similar significance (beta(`r H4_test_results$df.null`) = `r apa_num(H4_test_results$coefficients$mu[[2]])`, *p* = `r apa_p(summary(H4_test_results)[1]$coefficients$mu[2,4])`). This result is in line with hypothesis 4.

## Hypothesis 5
If a replication does not properly measure the variables of the original effect, the effect analysed within the replication may not be the same as the original. Establishing a link between QMPs and non-replicability should confirm suspicions of their detrimental effects on replications.

-   RQ5. What is the association between QMPs in replication studies and replication outcome?

Several earlier findings point towards a positive relation between successful replication and various types of good research practices. A similar relation is expected for good measurement practices.

-   H5. QMPs in replication research are negatively associated with replication of an original finding.

### Analysis
For hypothesis 5 a similar formula was used:

```{=tex}
\begin{equation}
g(QMP\_ratio_{i}) = \beta_{0} + \beta_{1} replicated_{i} (\#eq:H5Main)
\end{equation}
```

Where $replicated_{i}$ is a dummy coded variable, which was coded as 1 if the meta-analytic effect in the replication set matches the original effect in direction at a significance alpha level of .05, and 0 otherwise. Note here that unlike in the test for hypothesis 4, $QMP\_ratio_{i}$ refers only to the QMP ratios of the replications, meaning that original article QMPs were not included in this analysis.

```{r Hypothesis5Model, include = FALSE}
## Hypothesis 5 beta regression model
H5_test_results <- betareg(QMP_Ratio_data[QMP_Ratio_data$is_replication,]$QMP_ratio ~ coded_data_replications$hypothesis_support)

H5_test_results_REV <- betareg(QMP_Ratio_data[QMP_Ratio_data$is_replication,]$QMP_REV_ratio ~ coded_data_replications$hypothesis_support)

```

Results showed that successful replication significantly related to a decrease in the ratio of QMPs in replication protocols (beta(`r H5_test_results_REV$df.null`) = `r apa_num(H5_test_results_REV$coefficients$mu[[2]])`, *p* = `r apa_p(summary(H5_test_results)[1]$coefficients$mu[2,4])`). This result is in line with hypothesis 5. However, the results based on the initial QMPs did not show significant results (beta(`r H5_test_results$df.null`) = `r apa_num(H5_test_results$coefficients$mu[[2]])`, *p* = `r apa_p(summary(H5_test_results)[1]$coefficients$mu[2,4])`).

## Hypothesis 6
If essential information on a measure is not available in an original article, it may be difficult to know the specifics of that measure when used in a replication. In that case, the lack of proper reporting on measurement would impact the transparency and validity of future replications. To provide insight into this issue, an investigation of the relationship between original and replication QMPs was included.

-   RQ6. What is the association between QMPs in original research and QMPs in replications of psychological research?

If the relationship exists, it is expected that questionable measurement practices in an original paper might cause a spill-over effect of QMPs into the replication. Thus, QMPs are expected to be associated across original and replication studies.

-   H6. Total number of QMPs in original psychological research is positively related to total number of QMPs in replication research.

### Analysis
The relationship between original article QMPs and replication protocol QMPs was modelled to test hypothesis 6. The outcome variable remains the same as in the model for hypothesis 5. The independent variable is changed to the QMP ratio for original articles. The resulting model formula is as follows:

```{=tex}
\begin{equation}
g(QMP\_ratio\_replication_{i}) = \beta_{0} + \beta_{1} QMP\_ratio\_original_{i} (\#eq:H6Main)
\end{equation}
```

```{r Hypothesis6Model, include = FALSE}
## Hypothesis 6 beta regression model
H6_test_results <- betareg(QMP_Ratio_data[!QMP_Ratio_data$is_replication,]$QMP_ratio ~ QMP_Ratio_data[QMP_Ratio_data$is_replication,]$QMP_ratio)

H6_test_results_REV <- betareg(QMP_Ratio_data[!QMP_Ratio_data$is_replication,]$QMP_REV_ratio ~ QMP_Ratio_data[QMP_Ratio_data$is_replication,]$QMP_REV_ratio)

```

The results showed that total QMP ratio in the original article significantly predicted total QMP ratio in the subsequent replication (beta(`r H6_test_results_REV$df.null`) = `r apa_num(H6_test_results_REV$coefficients$mu[[2]])`, *p* = `r apa_p(summary(H6_test_results_REV)[1]$coefficients$mu[2,4])`), which provides some evidence in favor of hypothesis 6. However, this relation was not significant for the QMP ratios obtained with the initial protocol (beta(`r H6_test_results$df.null`) = `r apa_num(H6_test_results$coefficients$mu[[2]])`, *p* = `r apa_p(summary(H6_test_results)[1]$coefficients$mean[2,4])`).

## Pre-registration Analysis Justification & Reason for Omission

Follow-up analyses were preregistered for hypothesis 6. The goal of these analyses was to see if there are any type specific carry-over effects in QMPs between original and replication. Theses analyses were changed from the pre-registered analyses to beta-regressions for each unique combination of the QMP categories across original and replication, as the original analysis using contingency tables was later deemed improper for our data and hypothesis 6. Furthermore, these models no longer converged in this version of the article and so are therefore omitted. These analyses can be found in the first pre-print release of this project in the supplementary materials.