### 2025-10-06 SupplementaryMaterials/SupplementaryAnalysesScripts/Supplementary_Omega_analyses.rmd

<!-- Supplementary Analyses B -->

```{r analysis-preferences}
# Seed for random number generation
set.seed(17042023)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Supplementary Analyses D: Omega Results

This appendix contains

We additionally estimated McDonald’s Omega for each lab from the same set of replications as for Cronbach's Alpha, since it has been argued to be a more informative measure of reliability than Cronbach's Alpha, while also simultaneously providing some validity evidence [@crutzenScaleQualityAlpha2017b; @dengTestingDifferenceReliability2017]. McDonald’s Omega was calculated using the omega function in the *psych* R package [*v`r getNamespaceVersion("psych")[[1]]`*; @R-psych]. Default arguments were used in the function except the nfactors argument, which was set to 1. No RG Meta-Analysis was performed on McDonald's Omega, as we could not find a formula for calculating its standard error.

## Results

```{r check_on_Omega_convergence, warning = FALSE}
is_mostly_na <- function(x){
  return((sum(is.na(x)) / length(x)) > .5) 
}

Omega_convergence_vector <- tapply(calculated_alpha_plot_data$omega.tot, calculated_alpha_plot_data$g, is_mostly_na)

# sum(Omega_convergence_vector)
```

Figure \@ref(fig:PlotOmegaDistributions) below shows the result of our Omega analyses. Note that while we show the distributions of omega values here, due to the lack of a formula to calculate the Omega standard error, we did not perform a meta-analysis on these results. Overall, our Omega coefficient analyses show a similar pattern between measures to that observed for the alpha coefficient albeit with generally higher coefficient scores across measures. However, not all improved. The two columns beside Figure \@ref(fig:PlotOmegaDistributions) show in how many labs each measure was used, as well as for how many labs an omega coefficient could be obtained, which was not always. In `r apa_num(sum(is.na(calculated_alpha_plot_data$omega.tot)) / nrow(calculated_alpha_plot_data) * 100)`% of labs, no omega coefficient could be obtained. For `r sum(Omega_convergence_vector)` measures the number of labs for which an Omega coefficient could not be obtained was even the majority. This relates to our findings from the unidimensionality tests, where some factor models did not converge.

```{r PlotOmegaDistributions, warning = FALSE, fig.cap = "Distributions of calculated McDonald’s Omega t  calculated for the responses on a measure at each lab location, across the eighteen measures for which raw data was available from which McDonald’s Omega t coefficients could be calculated. The N.Omg column besides the figure shows the number of lab locations for which an Omega coefficient could be successfully calculated. Due to lack of underlying model convergence an Omega coefficient could not be calculated for all lab locations, or any lab locations for the Der Fruyt et al. (2000) measure. The N.tot column shows the total number of lab locations at which the measure was taken. Any difference between N.omg and N.tot can be seen as an indication of the instability to get a stable estimate of an underlying facto model for some lab locations for that measure."}
# plot for omega
ggplot(calculated_alpha_plot_data, aes(x = omega.tot, y = g)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(alpha = 0.1) +

  # adding labels to show the number of labs per measure for which a total omega could be calculated
  geom_text(data = calculated_alpha_plot_data[1:19,], label = table(calculated_alpha_plot_data$g[!is.na(calculated_alpha_plot_data$omega.tot)]), x = 1.15, y = 1:19, size = 2.8) +
  # adding labels to show the number of labs per measure
  geom_text(data = calculated_alpha_plot_data[1:19,], label = table(calculated_alpha_plot_data$g), x = 1.35, y = 1:19, size = 2.8) +

  # theming
  theme_minimal() +
  theme(legend.position = "none", plot.margin = unit(c(1, 6.5, 1, 1), 
                                                     "lines")) +

  # adding the replication line
  geom_hline(yintercept = 6.5, color = "red", size = 1) +

  # adding the necessary indicative texts
  annotation_custom(grob = textGrob(label = "Not Replicated", hjust = 0, gp = gpar(fontsize = 10)), ymin = 7.25, ymax = 7.25, xmin = 0.01, xmax = 0.01) +
  annotation_custom(grob = textGrob(label = "Replicated", hjust = 0, gp = gpar(fontsize = 10)), ymin = 6, ymax = 6, xmin = 0.01, xmax = 0.01) +
  annotation_custom(grob = textGrob(label = "N.Omg", hjust = 0, gp = gpar(fontsize = 12)), ymin = 20.2, ymax = 20.2, xmin = 1.08, xmax = 1.08) +
  annotation_custom(grob = textGrob(label = "N.Tot", hjust = 0, gp = gpar(fontsize = 12)), ymin = 20.2, ymax = 20.2, xmin = 1.28, xmax = 1.28) +

  coord_cartesian(xlim = c(0, 1), clip = "off") +

  ylab("") +
  xlab("Omega")

```